---
title: "Farms, fish, and cost: Locally-tailored objective functions expand utility of watershed model results and clarify trade-offs in an active local water planning process, northern California"
author: "Claire Kouba"
date: "2026-02-13"
output: 
  bookdown::word_document2:
    fig_caption: yes
    number_sections: true
    global_numbering: true
bibliography: library.bib

---

```{r setup1, include = FALSE}

# TO DO: 
# Add inset CA map to Fig 1
# Tie first sentence in conclusions to last sentence in intro
# Fix station numbers in caption for Fig 2

# Consider taking curly brackets off NOTICE OF APPROVAL info in .bib.


# ask thomas: how much of a scenario design supplement should there be?
# Snyder-Orloff work? citeable?
# Best way to to interpret grain ET results. Is grain more valuable than alfalfa? is there a fudge factor?


# old title: Multi-objective assessment of a stakeholder-defined portfolio of groundwater and stream management actions in an agricultural basin

knitr::opts_chunk$set(echo = TRUE)

# library(officedown) # for landscape pages?
library(sf)
library(terra)
library(geodata)
library(rpostgis)
library(tidyr)
library(plyr) #ssurgo and climate data
library(data.table) #fread function
library(dataRetrieval) # USGS flow data
library(MARSS)

library(lubridate)
library(httr)
library(stringr)
library(rstudioapi)
library(dataRetrieval) # for usgs data
library(data.table) #month function
library(zoo) # rolling means
library(rgl) # plot3d
library(RANN)
library(tmap)
library(RColorBrewer)
library(flextable)
library(scales) # for alpha() color function
library(ggthemes) # for colorblind_pal()
library(colorspace) # for sequential_hcl()
library(ggpubr) # to use get_palette()
library(here)


# Directories
# ms_dir <- dirname(getActiveDocumentContext()$path ) 
ms_dir = here::here()
data_dir = file.path(here::here(), "Data")
# wl_dir = file.path(data_dir, "Wells and water levels")
scratch_dir = file.path(here::here(), "scratch_work")
fig_dir = file.path(ms_dir, "Graphics and Supplements")
# Directory for postprocessed, tabulated results
svihm_results_dir = file.path(ms_dir, "Data", "SVIHM Model Results")
# Directories for SVIHM inputs and outputs, not yet postprocessed
svihm_scenarios_dir = "C:/Users/Claire/Documents/GitHub/SVIHM/Scenarios"
svihm_scenarios_backup_dir = "G:/SVIHM Scenarios Storage"

# Load spatial and tabular data
# If the data files don't exist locally,
if(!file.exists(file.path(ms_dir,"manuscript_data.RData"))){
  source(file.path(ms_dir,"01_DataRetrieval_Cleaning_SaveLocal.R")) #pulls data off server, local files, and internet
}
  
load(file.path(ms_dir, "manuscript_data.RData")) # Load data layers/tables
start_date = as.Date("1990-10-01")
INCLUDE_SUPP = F

```

```{r setup2, include = FALSE}

source(file.path(ms_dir,"02_Figure_Functions.R"))   # Load figure functions

# print figures for slides, yes or no?
save_figs_here = fig_dir#"C:/Users/Claire/Dropbox/Documents/UCD/Presentations_Talks_Workshops_miniprojects/2022.02.17 Peer Group presentation Ch. 1/figs"
print_for_slides=T; fig_i = 1
save_landuse_figure = F

```

# Abstract {-}

In the rural  American West, water needs of agriculture and ecosystems are often in competition. Furthermore, in places with interconnected surface- and groundwater systems, stream-aquifer exchange increases the difficulty of tracking impacts of diverse water users on each other and on the system as a whole. Hydrologic models can estimate the impacts of water uses and management actions on groundwater elevation and streamflow, but may provide overwhelming amounts of information and/or may be challenging for stakeholders to interpret. In this study, we address these challenges and aim to increase utility of hydrologic model results for a case study jurisdiction in northern California, which is in the process of implementing a long-term Groundwater Sustainability Plan (GSP). We developed a summary statistic or proxy for each of the three primary policy considerations in the GSP (environmental, agricultural, and project cost). We then used them to summarize the results of 25 management scenarios developed for the GSP, which were simulated using an existing integrated surface- and groundwater model. **We found that a trade-off in benefits for fish and farms was evident in every category of infrastructure investment (a proxy for project cost), though greater infrastructure investment can achieve some reductions in this trade-off. Additionally, although the GSP management priorities emphasized infrastructure investments, both infrastructure-based and regulatory approaches fell within the Pareto-optimal set of management options under a strict application of these objective functions.** Finally, **regarding regulatory actions, management interventions targeted at low-flow periods produced a more efficient gain in environmental flow value per cost in agricultural productivity.** This work demonstrates that basin-specific, stakeholder-informed objective functions can increase *stakeholder-salience of existing hydrologic model results*, and can help clarify trade-offs in ongoing local water planning efforts.


<!-- # LIT REVIEW -->

<!-- * [@SaleEtAlApproach1982]: reservoir op.  relat between bio instream flow needs and more traditional water project objectives. IFN performance objective: index of physical habitat conditions called weighted usable area (WUA) dev. by USFW. Hydraulic  method. "There exists a strong need for developing new methods which are capable of integrating ecological considerations into water resources decision-making" -->
<!-- * [@BartholowWaddleSearch1995]: options: "mgmt that will perform well despite possible states of nature" (over-engineer?) second option, adaptive management (Holling 1978). "For our purposes, variation in the freshwater environment may be broken into four principal factors: water supply, meteorology, spawner density, and spawner location". "The number of chinook migrating to the ocean is a function of the volume of water released (Bartholow and Waddle 1994)". complex population model simulates spawning, growth, mortality, and movement. input variables (spawner abundance and location), streamflow and water temp are weekly avg. Ran the stochastic model 1200 times. Generated an optimum flow regime. Modeled the relat. between flow and millions of outmigrants. "Besides adjusting flow on an annual basis to hydrologic, meteorologic, and biological variability, adaptive management could respond contingently to other event-driven phenomena, such as flushing the stream of built-up sediment." -->
<!-- * [@PaulsenWernstedtCostEffectiveness1995] "used simulation and optimization to minimize costs of many salmon recovery measures" -->
<!-- * [@CardwellEtAlDesigning1996] - Reservoir operation. fish population model. water supply shortage, habitat avail. for fish species during life stages. habitat capacity metric.  -->
<!-- * [@DraperEtAlEconomicEngineering2003]: CA major water supply economic-engineering optimization model (CALVIN?) Can suggest promising approaches for further study. "environmental objectives were omitted from optimization models or modelled as constraints of legally required minimum instream flows to remove them from economic valuation and decision making" -->
<!-- * [@JagerRoseDesigning2003] in Central Valley rivers, winter minimum flows.... Optimal flow regimes are also different for different species-specific environmental objectives, and vary based on water year type (e.g., a recommendation to provide high spring flows in wet years)  -->
<!-- * [@LabadieOptimal2004]: Water resource optimization methods review -->
<!-- * [@JenkinsEtAlOptimization2004]: Calvin, clearly. Increasing system flexibility (e.g. water transfers and exchanges) These changes reduce the cost of accomodating env. requirements. -->
<!-- * [@WatanabeEtAlEconomics2006]: localized restoration can achieve small temp reductions. big temp reductions make farther-away restoration more important. Maximizing fish production is different than maximizing stream length that achieves temp goals. -->
<!-- * [@JagerSmithSustainable2008] suggest that managers "conduct research to identify which features of flow variation are essential for river health and to quantify these relationships" -->

<!-- Relevant papers that cited NullLund2012: -->

<!-- * [@HorneEtAlOptimization2016]: review paper. we must improve adoption of optimization models among end users and stakeholders. We should talk to them during model development. -->
<!-- * [@AdamsEtAlEnvironmental2017] - Key finding: "Killing some fish early to preserve water in storage for later use effectively ‘‘hedges bets’’ for the worst-case low hydrologic conditions and can help to avoid later and larger fish kills."  -->

<!-- KEY SUMMARY: "Separate release targets for downstream ecological needs [Arthington, 2012] are generally based on: (a) habitat extent and suitability [Sale et al., 1982], (b) downstream modeled fish populations [Cardwell et al., 1996; Cioffi and Gallerano, 2012; Jager et al., 1997; Jager and Rose, 2003; Null and Lund, 2011], and/or (c) environmental goals based on specified hydraulic, hydrologic, water quality or political metrics [Tharme, 2003]" -->

<!-- "we define the benefit function to be the seasonal flow requirements of each life history stage of one keystone species of fish" but they didn't define the benefit function! "up to the experts" -->

<!-- "the dam operator is responsible for two cohorts over all hydrologic conditions: one developing cohort (e.g., salmon that survive growing from eggs laid to smolts migrating to the ocean; (cohort A) and a second cohort of adults returning after several years at sea to spawn (cohort B)." -->

<!-- In their example, "The hydrologic needs of the fall-run Chinook track the natural hydrograph" -->
<!-- Trade-off between expected final storage and fish survival (diff for 2 cohorts) -->

<!-- System performance was assessed  -->

<!-- * [@BayazitUnalEffects1990] - rel, res, and mean and max deficit. Standard Operating Policy is best for rel. and res, and fine for deficits. Hedging improves mean deficit performance. Hedging can mean forsaking some reliability. -->
<!-- * [@YangCaiReservoir2011] (multiobjective framework) - "minimizes flood damage ... and maximizes fish diversity". Adding eco objective "can improve downstream fish habitat without jeopardizing its original flood control objective" -->
<!-- * [@ChenOldenDesigning2017] - relate hydro regimes and fish species abundance (native and nonnative). Multi-objective optimization. Balanced human water needs with helping native fish and killing nonnative. Many opportunities to advance all objectives - better than achieved by natural flow mimicry. "designer flows can offer multiple ecological and societal benefits in human-altered river" -->
<!-- * [@KraftEtAlPrioritizing2019] - two objectives. prioritize dam removal by 1) maximize connected aquatic habitat and 2) minimize water scarcity.  -->
<!-- * [@ZamaniSabziEtAlStrategic2019] - studying water conservation incentives in the Red River. spatial palnning framework. "First, even optimal allocation of water across the network cannot simultaneously meet societal and environmental flow targets given current water availability and societal usage. Second, modest reductions in societal water usage (e.g., 1–3%) can significantly reduce water conflicts in the basin, but only when these water conservation actions are strategically allocated in time and space. Third, however, substantial reduction in societal water usage (> 21%) is needed to fully satisfy societal and environmental flow goals."  -->
<!-- * [@FarzanehEtAlStochastic2021] "We quantify tradeoffs between human and environmental water needs and identify water sustainability strategies that are resilient to stochastic drought events" Red River south-central US. "Mathematical model for selecting the best water conservation policy." Harsher trade-offs between society and ecological objectives when you want higher reliability. (didn't work with stakeholders per se.) -->
<!-- * [@SanzEtAlProposal2025] Test e-flow regime. hydraulic modeling and habitat suitability. Tagus river in spain.  -->

<!-- [@BorkHirokawaTrends2021] - calls for more research into adaptive management and ecosystem services trade-offs. Better coordinate scientific research and knowledge needs of local gov -->
<!-- [@NordgrenEtAlSupporting2016] - "Local governments are on the front line of efforts to address climate-related impacts" "Findings demonstrate that existing services and resources are meeting the early phases of local adaptation efforts such as conducting vulnerability assessments and creating adaptation plans, but are failing to meet the needs associated with implementing, monitoring, and evaluating adaptation activities". streamline resource access.  -->

\newpage

# Introduction

Climate adaptation in the water sector, in its many facets - e.g., securing water and food supplies, updating infrastructure to reduce flood and drought risk, and preserving ecosystem function - is often most effective when implemented at the local level [e.g., @VanSteenbergenPromoting2006]. However, local water managers face a complex regulatory landscape and serve diverse stakeholders, and typically have unmet knowledge needs [@BorkHirokawaTrends2021; @NordgrenEtAlSupporting2016]. Academic hydrologists have many tools to fill such knowledge gaps; however, when seeking broadly-relevant findings, they commonly avoid time-consuming engagement with specific local water organizations, which can reduce the utility of the resulting research to the relevant decisionmakers [@FreemanWICKED2000]. In this study, we take advantage of the authors' involvement with an existing water resource planning process, and a previously-developed integrated surface- and groundwater model, to test the utility of introducing quantitative objective functions into the public prioritization of groundwater management alternatives in a case study jurisdiction.

<!-- Even a dedicated case study can be useless unless designed in discussion with stakeholder priorities. -->
<!-- Sustainable development of water resources, in an era of climate change and increasing human water consumption [e.g., @Williams2020; @Rodell2018], requires an interdisciplinary understanding of connected socio-hydrologic systems [@Reuss1992; @DiBaldassarre2019]. One such interdisciplinary approach is water resources systems analysis (WRSA), which involves applying optimization methods to water resource planning [@Maass1962; @CohonMarksReview1975; @Brown2015]. The utility of such studies typically depends on robust definitions of the "objective function(s)", which summarize and quantify the desired outcome(s) of the proposed project or management scenario [@Gorelick1983]. In practice, classical WRSA approaches have been criticized as being too removed from stakeholder concerns [e.g., @Rogers1986a].  -->

In our study area, the intermontane rural Scott Valley in northern California (Figure \@ref(fig:scottLanduseMap)), the local government and public stakeholders spent three years (2018-2021) developing a plan for the long-term use of groundwater (a Groundwater Sustainability Plan or GSP). In that process, they evaluated dozens of potential water management scenarios [Siskiyou County -@SiskiyouCountyScott2021]. In public meetings, different objectives were often championed by individual stakeholders or interest groups. An integrated hydrologic model, the Scott Valley Integrated Hydrologic Model, or SVIHM [@TolleyEtAlSensitivity2019], was available to support this policy-making process. The numerical streamflow results of simulated management scenarios drove decision-making, but discussions on the balancing of multiple objectives were typically qualitative. In effect, the optimization framework used to arrive at the ultimate list of management priorities was implicit, and the conceptual weight given to different objectives varied between stakeholders.

Differences between stakeholders in what is considered an acceptable trade-off are common in policy decisions [@MonarchiEtAlInteractive1973]. Consequently, multi-objective optimization studies often avoid proposing a single optimal solution; instead they identify a Pareto-optimal or non-inferior set of options, which can clarify the unavoidable trade-offs in the performance of various system objectives [e.g., @CohonMarksMultiobjective1973]. Methods for finding the non-inferior set often involve uniform or randomized initial sampling of the decision space [e.g., @MonarchiEtAlInteractive1973; @YehBeckerMultiobjective1982; @KapelanEtAlMultiobjective2005]. In the case of the Scott Valley GSP, the management actions considered by stakeholders covered a wide range of decision variables (e.g., land use changes in a variety of spatial configurations, technology and/or infrastructure investment, and regulatory interventions on a range of growing season dates). Because a full randomized sampling of this decision space would be computationally prohibitive, this study instead focused on a set of simulated management scenarios that were effectively pre-screened by stakeholders for feasibility in the landscape of existing land uses and political dynamics. 

In this study we developed numerical representations of three objectives: 1) maximizing specific environmental flows, 2) minimizing reduction in agricultural water supplies, and 3) minimizing new infrastructure costs. With these, we quantitatively summarized the management benefits and trade-offs within the existing suite of several dozen stakeholder-proposed management scenarios. We hypothesized that an explicit objective function framework would yield different prioritization of management actions than the set agreed upon in public meetings (as recorded in the 2021 GSP) and potentially new insights about the holistic management option landscape. 
<!-- This work could help frame ongoing water planning efforts in Scott Valley, and constitutes an approach for supporting locally-driven water planning efforts in California and elsewhere. **possibly remove final sentence** -->

```{r scottLanduseMap, echo = FALSE, fig.cap = "The Scott River flows south to north through Scott Valley before flowing through a canyon and joining the Klamath River. Land uses in the cultivated areas of the Scott River watershed are shown (adapted from the 2016 DWR Land Use Survey), as well as the location of several communities and the Fort Jones streamflow gauge (USGS Station ID 11519500). \\label{fig:scottLanduseMap}" , fig.height = 8.5, warning = F}

lu_graphic_filename = file.path(fig_dir,
                                "Ch. 2 Scott valley setting and land use.png")

if(save_landuse_figure == T){
  lu_tm = landuse_figure_2016()
  tmap_save(tm = lu_tm,
            filename = lu_graphic_filename,
            width = 7, height = 8.5, units= "in", dpi = 300)
}

knitr::include_graphics(lu_graphic_filename)
```


# Background: Case study and GSP development

## Hydrography and climate

The Scott River drains a 2,109 km^2^ (814 square mile) intermontane watershed known as Scott Valley, with forested uplands and a largely agricultural valley floor (Figure \@ref(fig:scottLanduseMap)) overlying an alluvial aquifer up to 79 m (260 ft) thick [SWRCB -@CaliforniaStateWaterResourcesControlBoardSWRCBReport1975; @MackGeology1958]. A key stream gauge (USGS Station ID 11519500, or the Fort Jones Gauge or FJ Gauge) is located downstream of the watershed's cultivated areas (Figure \@ref(fig:scottLanduseMap)).

<!-- used to assess hydrologic conditions and inform water management decisions in the populated areas of the valley [e.g., Siskiyou County -@SiskiyouCountyScott2021]. -->

<!-- . The valley floor overlies unconsolidated alluvial sediments of up to 79 m (260 ft) thickness [SWRCB -@SWRCB1975]. The sediments form a mostly unconfined aquifer system bordered by low permeable, fractured underlying and adjacent bedrock composed of mostly Paleozoic metasedimentary formations [@MackGeology1958].  -->

Scott Valley has a Mediterranean precipitation regime with distinctive seasons of cold, wet winters and warm, dry summers (Figure \@ref(fig:precipAndTempFigure), panels A and B). Winter storms (October-April) and melting mountain snowpack (April-July) generate streamflow during the wet and growing seasons. As a result, the Scott River and tributary streams experience seasonal high and low flows (Figure \@ref(fig:precipAndTempFigure), panel C).

Streams and groundwater are highly interconnected in the Scott River watershed [@TolleyEtAlSensitivity2019]. Tributary streams and the Scott River are a source of recharge to the aquifer, and groundwater discharge sustains streamflow in some areas, especially during the dry season of August-October or November. Some streams are ephemeral, i.e., some tributaries or reaches of the Scott River go dry in average-to-dry water years [@TolleyEtAlSensitivity2019].


```{r precipAndTempFigure, echo=FALSE, fig.cap="\\label{fig:precipAndTempFigure} Summary of temperature (Panel A), rainfall and reference evapotranspiration (Panel B), and daily streamflow values (Panel C) in Scott Valley. For precipitation, regression relationships between the Callahan and Fort Jones weather station records and the records of NOAA stations in Greenview, Etna, and Yreka (USC00042899, USC00043614, USC00049866, and US1CASK0005) were used to predict missing values (i.e., fill gaps) in the Callahan and Fort Jones records; the daily mean of the two stations was used as the final precipitation record for purposes of this figure and the SVIHM precipitation input." , fig.height=8.4, fig.width=7, warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))


if(print_for_slides==T){
  png(filename = fig_path, width = 8.4, height = 7, units = "in", res = 300)
  precip_temp_flow_figure()
  dev.off()
}

knitr::include_graphics(fig_path)
fig_i = fig_i +1

```

## Land and water uses

<!-- The population of Scott Valley is approximately 5,200, based on area-weighted block-level 2020 census data, including the populations of the two incorporated towns of Etna and Fort Jones [population 678 and 695, respectively; @USCensus2021].  -->

Because the watershed is undammed, managers and water users influence Scott River flow primarily via diversion of surface waters or pumping of groundwater. Consequently the most powerful tool available to manage Scott River water flow is regulation of land use and thus water demand [Siskiyou County -@SiskiyouCountyScott2021].

About two thirds of the land within the Scott River watershed is under private ownership with the remaining area managed by the Quartz Valley Indian Reservation, the United States (U.S.) Department of the Interior Bureau of Land Management (BLM) and U.S. Forest Service (USFS) [@HarterHinesSCOTT2008]. On the valley floor, according to land use surveys conducted by the California Department of Water Resources [DWR -@CaliforniaDepartmentofWaterResourcesStatewide2017], half of the area overlying the alluvial aquifer is covered by irrigated agriculture, notably pasture for cattle and an alfalfa/grain rotation (Figure \@ref(fig:scottLanduseMap); Table \@ref(tab:scottLanduseTable)).

Historically, local regulation of land use has focused on maintaining the rural and agricultural character of Scott Valley [@ScottValleyAreaPlanCommitteeScott1980]. However, during recent declarations of emergency drought conditions, water use has been periodically curtailed to preserve flows for aquatic species [SWRCB -@StateWaterResourcesControlBoardEstablishment2021a; -@StateWaterResourcesControlBoardEstablishment2024a; -@STATEWATERRESOURCESCONTROLBOARDRESOLUTION2024].

<!-- [SWRCB -@StateWaterResourcesControlBoardEstablishment2021; SWRCB -@StateWaterResourcesControlBoardEstablishment2024; SWRCB -@STATEWATERRESOURCESCONTROLBOARDRESOLUTION2024]. -->

<!-- Regulating land use to improve ecological outcomes entails significant economic, political and social risks, because much of the economic activity in this area is related to agriculture. In addition, Scott River conditions influence fish population dynamics both within the watershed and in the broader Klamath system.  -->

<!-- Fort Jones and Etna are categorized as severely economically disadvantaged (defined as communities with an annual median household income [MHI] of less than 60% of the average annual MHI in California). Based on the 2013-2017 American Community Survey Five Year Estimates, the statewide annual MHI is \$67,169, and Fort Jones and Etna both qualify as SDACs with annual MHIs of \$29,662 and \$35,333, respectively [@USCensusBureau2018]. The unincorporated community of Greenview is listed in government databases as a disadvantaged community (defined as an MHI of less than 80% of average state MHI), but no MHI data are available for this community [@DWR2019]. -->

<!-- The health of the Klamath salmon run has implications for commercial fishing, recreational activities, and cultural practices of Native American tribes in the region, including the Quartz Valley Indian Community and the Karuk and Yurok Tribes [@Graham2012]. -->


<!-- All of the regulatory and management programs in this region, including recommended instream flows [CDFW -@CaliforniaDepartmentofFishandWildlifeINTERIM2017] and legal rights governing surface water diversion [@CaliforniaStateWaterResourcesControlBoardScott1980], are tabulated in units of cubic feet per second (cfs). For consistency, this document will also use primarily cfs units, with metric units in parentheses." -->

<!-- Like many rural watersheds in arid to semi-arid regions, balancing environmental and agricultural water demands is a significant challenge faced by local water managers. Water uses, land management, and the measurement units used in local regulatory documents were discussed in Chapter 1 of this dissertation:  -->

<!-- > "Water in Scott Valley is used for agricultural, domestic, and municipal supply. It also facilitates recreation and provides Native American cultural services, among other designated beneficial uses [NCRWQCB -@NCRWQCB2006b]. -->






```{r scottLanduseTable, echo = FALSE, results = 'asis'}

calc_landuse_table = F

if(calc_landuse_table == T){
  lu_tab = landuse_table_2016()
  write.csv(lu_tab, file.path(data_dir,"land use table 2016.csv"), quote = F, row.names = F)
}
lu_tab = read.csv(file.path(data_dir,"land use table 2016.csv"))

#Create caption and column labels
caption_text = "Acreage and percent of total Basin area covered by generalized land uses as reported in DWR's 2016 land use survey."
column_labels = c("Land Use Category", "Acres", "Percent of Watershed Area")
colnames(lu_tab)=column_labels

#declare xtable object, assign formatted column labels, and print to latex object
# landuse_table = xtable(lu_tab, label = "tab:scottLanduseTable",
#                  caption = caption_text, digits = c(NA, NA, 0, 1))
# names(landuse_table) = column_labels
# print.xtable(landuse_table, comment = F,type = "latex", format.args = list(big.mark = ","), include.rownames = F)
#caption.placement = "top",


lu_tab_ft = flextable(data = lu_tab)
lu_tab_ft = set_caption(lu_tab_ft, caption = caption_text)
lu_tab_ft = autofit(lu_tab_ft)
lu_tab_ft

```



### Environmental stream flows

Sufficient environmental flows, varying by season, are necessary to the survival of vulnerable species in the Scott River watershed [@ScottRiverWatershedCouncilRestoring2018]. Coho salmon (*Oncorhynchus kisutch*) in Scott Valley belong to the Southern Oregon / Northern California Coast (SONCC) Evolutionarily Significant Unit (ESU), and were listed as threatened under the federal and California Endangered Species Acts (ESAs) in 1997 and 2005, respectively. Statewide coho populations have declined by an estimated >90% since the 1940s [@BrownEtAlHistorical1994].

To complete their life cycle, coho salmon require several key functional flows in the Scott Valley freshwater stream system [@KoubaEtAlWatershedSpecific2025], including the following:

* Fall flows, increasing after the dry season in response to early wet season storms, that are sufficient to provide coho spawners access to preferred spawning reaches within the fall-run time window;
* Flows providing habitat for rearing juvenile salmon for their ~12-15 months of freshwater residence; and
* Sufficient flow in the spring for yearling smolts to migrate to the ocean.

While Chinook also require fall flows for spawning, their specific flow requirements are somewhat distinct from coho due to their different life histories:

* Chinook juveniles spend less than one year in the freshwater environment [@AgrawalEtAlPREDICTING2005], while coho overwinter and typically migrate to the ocean at 1+ years old [@MoyleCoho2002; @McMahonHabitat1983]; 
* Chinook tend to spawn earlier in the fall than coho [@KnechtleGiudice20192020; @MagranetScott2015a; @MagranetScott2017];
* Chinook prefer to spawn in the main stem of the Scott River, while coho prefer to spawn in tributaries [e.g., @MagranetScott2018; @MagranetYokelScott2017], making juvenile Chinook salmon more vulnerable than coho to egg burial or bed scour during high-flow storm conditions.

Different functional flows are needed by other species, such as cottonwood trees, which require flood-stage flows to disperse their seeds [@MahoneyRoodStreamflow1998]; but in this work we focus on coho and Chinook salmon as two examples in developing a single-species environmental objective function. Alternative objective function formulations are discussed in later sections.


### Agricultural water demand 

Agricultural water demand in Scott Valley has been **investigated?** **Cite Orloff/Snyder work?** simulated for a historical model period of Oct. 1, 1990 to Sept. 30, 2025, using the calibrated Scott Valley Integrated Hydrologic Model (SVIHM) [@HarterHinesSCOTT2008; @FogliaEtAlScott2013; @TolleyEtAlSensitivity2019].  Water demand for crops and native vegetation (evapotranspiration, or ET, in Figure \@ref(fig:avgWaterBudgetFig)) and applied irrigation volumes (diverted surface water and pumped groundwater, noted as SW and GW Irrigation in Figure \@ref(fig:avgWaterBudgetFig)) are highest in the late summer. Annual crop ET and irrigation demand is relatively stable year to year while precipitation, and thus recharge to groundwater, are more variable (Figure \@ref(fig:annualSWBMBudget)).

In this intermontane valley, the growing season is constrained by hard frosts, and begins in March-April and extends through September-October. Both of the primary crops (alfalfa and pasture) are perennial, and will continue producing through the growing season as long as soil moisture is available. Alfalfa cultivation since the 1970s has commonly involved three to four harvests (referred to as cuttings) per growing season; it is typically irrigated several times and then left to dry before cutting [@TolleyEtAlSensitivity2019].

<!-- In the mid-twentieth century, farms and ranches irrigated grain, alfalfa and pasture crops mainly with diverted surface water, which was abundant until mid or late summer, and many utilized flood irrigation. On this schedule two alfalfa cuttings were harvested, with the last irrigation typically occurring in July. Since the 1970s, widespread construction of irrigation wells has effectively extended the irrigation season and allowed alfalfa farmers to harvest three or four cuttings in a growing season. (Additionally, higher-irrigation efficiency technology like wheel-line and center-pivot sprinklers became common, though this does not alter the consumptive demand of the cultivated crops.) Based on historical crop coverage and water demand estimates by Harter and Hines [-@HarterHinesSCOTT2008], "the amount of water likely used by crops has increased from 1958 to 2000 by between 15 percent (10,000 more acre feet) and 30 percent (20,000 acre feet) depending on the date when surface irrigation [is assumed to have ceased historically], i.e. July 15, Aug 1 or Aug 15." -->

The tension between agricultural and ecological water demands is highest at the end of the dry season, when groundwater elevations and surface flows are at their lowest, crop demand remains high, and anadromous fish require access to spawning habitat. The management of these fall flows was a central concern of GSP development in 2018-2021.

```{r avgWaterBudgetFig, echo=FALSE, fig.cap="\\label{fig:avgWaterBudgetFig} Average monthly water budget for the soil zone, simulated with SVIHM, for the model simulation period (water years 1991-2018). Positive and negative values indicate average net monthly fluxes into and out of the soil zone, respectively." , fig.height=3.8, fig.width=7, warning=F}


fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))

if(print_for_slides==T){
  avg_swbm = avg_swbm_budget_fig(scen_id = "basecase")
  ggsave(filename = fig_path, plot = avg_swbm, width = 7, height = 4, units = "in", dpi=300)
}

knitr::include_graphics(fig_path)
fig_i = fig_i +1


```


```{r annualSWBMBudget, echo = F, warning =F, fig.cap="\\label{fig:annualSWBMBudget} Annual values for inflowing (positive) and outflowing (negative) water budget components in the SVIHM soil zone (historical basecase simulation). Irrigation refers to both groundwater and surface water sources; Recharge refers to water percolating through the soil column and recharging the aquifer below. Both crop ET and ET from native vegetation remain relatively stable year to year, while precipitation inputs, and thus recharge to the aquifer, have higher interannual variability." , fig.height=3.8, fig.width=7}


fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))

if(print_for_slides==T){
  png(filename = fig_path, width = 7, height = 4, units = "in", res = 300)
  annual_swbm_budget_fig(scen_id = "basecase")
  dev.off()
}

knitr::include_graphics(fig_path)
fig_i = fig_i +1




```


``` {r calc_et_depth, warning = F, echo = F}
# calculate average annual ET depth fromcrops vs natveg
nv_area = sum(st_area(svihm_fields[svihm_fields$landuse_desc=="ET/No Irrigation",]))

crop_area = sum(st_area(svihm_fields[svihm_fields$landuse_desc %in%
                                       c("Pasture","Alfalfa/Grain"),]))

# pull water budget values


swbm_monthly = get_swbm_tab_with_separated_et(scen_id = "basecase")
swbm_monthly$Month=NULL

# # check sums
# swbm_monthly$et_crop_plus_nv = swbm_monthly$et_crop + swbm_monthly$et_natveg
# swbm_monthly$diff = swbm_monthly$ET + swbm_monthly$et_crop_plus_nv

# aggregate to annual
swbm_annual = aggregate(.~Water_Year, swbm_monthly, function(x){sum(x) / 1E6})

et_nv_mean = mean(swbm_annual$`Nat. Veg. ET`)
et_crop_mean = mean(swbm_annual$`Crop ET`)

et_nv_depth_m = et_nv_mean * 10^6 / nv_area * -1
et_crop_depth_m = et_crop_mean * 10^6 / crop_area * -1
```

## GSP development

The Scott Valley Groundwater Sustainability Plan (GSP) was developed by the Siskiyou County Flood Control and Water Conservation District, which is the designated Groundwater Sustainability Agency (GSA) charged by the state of California to sustainably manage local groundwater resources pursuant to the Sustainable Groundwater Management Act (SGMA) of 2014. GSP development took place between 2018 and 2021, and was guided by discussions in the monthly public meetings of an Advisory Committee composed of nine local stakeholders. Technical analyses supporting the public meeting deliberations and GSP policies were provided by several technical experts, including the authors of this study [Siskiyou County -@SiskiyouCountyScott2021].

SGMA requires each local GSA to define sustainability in the negative, by quantifying and then avoiding "undesirable results" that would be "significant and unreasonable". The central undesirable result in Scott Valley is streamflow depletion due to groundwater pumping. Consequently, a central challenge of GSP development was deciding the amount of (estimated) streamflow depletion that the GSA should aim to reverse, accounting for environmental water demand and what degree of water security was owed to local agriculture.

During GSP development, the authors of this study simulated dozens of land and water use scenarios to facilitate the comparison of management alternatives that could reduce streamflow depletion to various degrees. Simulations were conducted with a spatially explicit, integrated surface water-groundwater model [@TolleyEtAlSensitivity2019], as described below in Section 3.3.

# Methods

After 2021, we expanded on this public planning effort with a research study, based on the suite of management scenarios designed for the GSP, in the hopes that it could inform water management decisions during 5-year GSP updates over the coming decades. 

In particular, we developed two quantitative objective functions and one categorical objective proxy to evaluate the management scenarios developed for the GSP. The three objectives were: 

1) maximize specific environmental flows, 
2) maintain current levels of agricultural productivity (represented as minimizing reduction in agricultural water use), and 
3) minimize new infrastructure costs (represented as a categorical infrastructure/feasibility proxy) (Table \@ref(tab:objFunctionTable)). 

We then applied these to the results of each simulated management scenario, found the Pareto-optimal set of scenarios, and compared them to the prioritization of management actions adopted in the GSP [Siskiyou County -@SiskiyouCountyScott2021]. 

<!-- We should note that protecting domestic water use was also discussed as a management objective, but this was not evaluated as an objective function in management scenarios, for three main reasons: it was assumed that solving the environmental flows problem would intrinsically keep groundwater levels sufficiently high to protect domestic groundwater use from well outages; the number of reported well outages in recent dry years was minimal; and the data necessary to assess domestic well outages (specifically, total well depth and pump depth information) was available only for a small number of wells.  -->


```{r objFunctionTable, echo = FALSE, results = 'asis'}


#Create caption and column labels
caption_text = "Management objectives and description of objective functions applied to summarize scenario performance."
column_labels = c("Management Objective", "Objective Function")

mo_tab = data.frame(mgmt_o = c("Maximize specific environmental flows", 
                    "Minimize reduction of agricultural water use", 
                    "Minimize new infrastructure investment"),
                    obj_f = c("Hydrologic Benefit function, based on flow metrics empirically correlated with coho salmon reproduction rates (higher values preferred)",
                              "Crop ET volume over model domain (higher values preferred)",
                              "Feasibility category of 1-4 (lower values preferred)"))
colnames(mo_tab) = column_labels

mo_tab_ft = flextable(data = mo_tab)
mo_tab_ft = set_caption(mo_tab_ft, caption = caption_text)
mo_tab_ft = flextable::width(mo_tab_ft,j=1,  width = 2.5)
mo_tab_ft = flextable::width(mo_tab_ft,j=2,  width = 3)
mo_tab_ft

```


## Objective functions

### Hydrologic Benefit function

```{r read_HB_coefs, echo = F}

coefs_tab = read.csv(file.path(data_dir, "kouba_et_al_2025_ensemble_coefs_chinook and coho.csv"))
model_names = colnames(coefs_tab)[2:13]
capmaker_all_models = model_name_to_coefs_df(model_names)

```

Conceptually, the Hydrologic Benefit (HB) function aims to answer the question: to what degree does this water year provide enough flow, at the right time, for this species?

In a previous study, we empirically identified hydrograph features of high importance to Scott River coho and Chinook, based on correlations between hydrology and ecological observations such as outmigrating juvenile abundance [@KoubaEtAlWatershedSpecific2025]. The HB function is represented as a linear combination of these high-importance annual functional flow values. 

Functional flows are set of streamflow metrics related to the timing, amplitude, and duration of flow phenomena [@YarnellEtAlFunctional2015; @PattersonEtAlHydrologic2020; @CarpenterAccurately2024]. In the HB function, these were weighted according to their power to predict coho and Chinook outcomes, such as smolt production or smolt production normalized per female spawner. The combination of these weighted metrics can be interpreted as a prediction of coho salmon reproduction *supported by hydrologic factors*. Importantly, the HB function selected for this exercise only explains about *40% to 60%* of variation in observed smolt abundance, and the remaining 60%-40% of variation is assumed related to other life conditions, e.g., number of returning spawners, water quality, food resources, and internal fishery population dynamics [@KoubaEtAlWatershedSpecific2025]. 

There are many different possible HB function formulations; one example HB function is included below and three additional formulations are explored in Supplemental Material. These four HB functions correspond to the two species (coho and Chinook) as measured using two different units of annual salmon reproduction (outmigrating smolt abundance and normalized smolt-per-adult-spawner in the relevant brood year). See Supplemental Material for more HB function design information.

The example HB function highlighted here is based on coho juvenile outmigrant abundance. It predicts that a higher number of coho smolt produced (coho smolt-equivalent) is related to higher fall flows during a cohort's parents' spawning, lower fall flows during a cohort's time as overwintering yearlings, and a slower spring recession (see coefficient values in Table \@ref(tab:coefTabMainText)).


```{r coefTabMainText, echo = FALSE, results = 'asis'}

model_name = "coho_lasso_juv_on_hydro"
# model_name = "coho_lasso_jps_on_hydro"
# model_name = "chinook_lasso_juv_on_hydro"
# model_name = "chinook_lasso_jps_on_hydro"


coef_tab = coefs_tab[,c("Predictor",model_name, "Greater.Hydrologic.Benefit.value.associated.with")]
coef_tab = coef_tab[coef_tab[,model_name] != 0,]

#Create caption and column labels
attr = model_name_to_coefs_df(model_name)
caption_text = results8_cap_maker(attr)
column_labels = c("Predictor", "Value", "Greater Hydrologic Benefit value associated with")

colnames(coef_tab) = column_labels

hb_ft = flextable(data = coef_tab)
hb_ft = set_caption(hb_ft, caption = caption_text)
hb_ft = flextable::width(hb_ft, j = 1,  width = 1.5)
hb_ft = flextable::width(hb_ft, j= 2,  width = 1)
hb_ft = flextable::width(hb_ft, j=3,  width = 3)
hb_ft

```

Functional flow metrics informing the HB function were calculated from the observed daily average flow record of the FJ gauge ($FJ_{obs}$), using signal processing algorithms developed by the California Environmental Flows Framework [@YarnellEtAlFunctional2015; @PattersonEtAlHydrologic2020; @CarpenterAccurately2024]. One key model result of the historical basecase scenario in SVIHM is a *simulated* estimate of the historical FJ gauge record ($FJ_{sim,~basecase}$). However, although SVIHM was extensively calibrated [@TolleyEtAlSensitivity2019], because all models involve structural simplifications of a full-scale watershed, $FJ_{sim,~basecase}$ sometimes over- or under-estimates $FJ_{obs}$. To control for this, all basecase HB values used in this study are derived from $FJ_{sim,~basecase}$, making them directly comparable to HB values derived from the simulated FJ flow records in the hypothetical management scenarios.
<!-- As a result of these differences, HB values calculated from flow metrics of $FJ_{sim,~basecase}$ tend to be higher than HB values derived from $FJ_{obs}$ by a mean of **4.5** coho spf-equiv. -->

### ET volume function (agricultural benefit proxy)

The annual crop ET volume, which is explicitly simulated in the hydrologic model, is used here as a proxy for agricultural productivity. Other options for framing the agricultural water use objective function include crop yields and total revenue, which may be of greater concern to stakeholders; however, annual crop ET was selected because the impacts of reduced water availability on yield or revenue will likely vary by grower and over time.

<!-- Specifically, conversion of ET to yield or revenue would introduce multiple uncertainties, such as those related to marginal crop production costs and available technology [@Howitt1995]. In studies that do implement a water-to-yield or -revenue conversion, multi-year average values for per-acre yield and/or revenue are commonly used [e.g., @Cole2021 for the Scott Valley case], since those are typically the only values available. This approach is most valid when the marginal change in productivity is small relative to the average [@Howitt1995]. Additionally, average values cannot account for time-varying costs of a management action [e.g., the increase in cost of lettuce cultivation due to banning a certain pesticide is higher in some seasons than in others, as described in @SundingMeasuring1996]. -->

Additionally, for alfalfa and pasture, which continue growing as long as water is available, a strong relationship between crop ET and yield is assumed [@CarterSheafferAlfalfa1983]. The assumption that crop ET is consistently related to yield or revenue breaks down for a management scenario that includes changing crops; this is explored further in the Discussion section. But on the merits of these options, annual crop ET was selected to represent agricultural benefit. 

We assume that grain, another important local crop, has a different relationship than alfalfa between ET and revenue, as grain is harvested once per year in mid-to-late summer and alfalfa is harvested three to four times per year in this region [@HarterHinesSCOTT2008]. This makes interpretation of the agricultural objective functions more complicated for three crop change scenarios in which alfalfa is converted to grain; see additional discussion in Section 5.

### Proxy for scenario infrastructure investment

In lieu of detailed cost estimates or feasibility studies, we chose to group management alternatives in qualitative categories approximating the degree of new infrastructure investment needed in each alternative. The categories describing project feasibility and timeframe are as follows:

0. No feasibility barriers (basecase; maintain current practices)
1. Implementable with current infrastructure 
2. Requires new infrastructure that would take 1-2 years to build
3. Requires new infrastructure that would take 3-5 years to build
4. Requires new infrastructure that would take 5+ years to build

This reflects the first-order understanding that, for example, building a reservoir (assigned to category 4) will take more effort and resources in terms of permitting and construction than widespread installation of higher-efficiency irrigation technology (assigned to category 2). See Table \@ref(tab:scenarioResultsTabMainText) for full list of assigned categories.

## Analyses of objective function values

### Trade-off efficiency

Combining two of the three objectives, scenarios can also be compared in terms of environmental benefit per agricultural cost. This "trade-off efficiency" of a management scenario, or $Eff~scen~$, is unitless (i.e., it has units of smolt-equiv. per smolt-equiv. / million m^3^ per million m^3^) (see equations below).

<!-- As discussed previously, balancing agricultural and environmental water demands in this watershed typically takes the form of seeking management actions that can improve flow conditions while minimizing the resulting reduction in agricultural water availability; in simple terms, to help fish while not harming farms. These objectives are non-commensurable (i.e., an estimated level of fish reproduction cannot be converted into units of agricultural productivity), making an explicit quantification and consideration of benefit trade-offs a critical feature of this decision-support analysis [@CohonMarksReview1975]. Environmental benefits are less commonly quantified than agricultural costs [e.g., @SundingMeasuring1996, which assessed numerical costs of a management action and assumed qualitative environmental/human health benefits], but the HB function makes this trade-off quantification possible in the Scott Valley case.  -->

<!-- For this analysis, which is not intended to identify one "best" or optimal solution, these trade-offs can be considered using scenario objective value differences relative to their basecase (historical simulated) values. We will call this quantity the "trade-off efficiency" of a management scenario. For a given scenario, the trade-off efficiency, or $Eff~scen~$, is unitless (i.e., it has units of smolt-equiv. per smolt-equiv. / million m^3^ per million m^3^) (Equation \@ref(eq:hbAvg)). -->

$$HB_{avg} = \frac{1}{n_{wy}}\sum_{i = wy_1}^{wy_n}{HB_{wy_{i}}}$$

$$ET_{avg} = \frac{1}{n_{wy}}\sum_{i = wy_1}^{wy_n}{Crop~ET_{annual,~wy_{i}}}$$

$$Eff_{scen} = \frac{(HB_{avg,~scen}-HB_{avg,~basecase}) / HB_{avg,~basecase}}{-1 * (ET_{avg,~scen}-ET_{avg,~basecase}) / ET_{avg,~basecase}} $$


### Pareto-optimal set of management actions

After calculating the objective function values for each scenario, we applied a simple algorithm to identify the Pareto-optimal set of scenarios [@ParetoCours1896]. A scenario was identified as Pareto-optimal if there were no other scenarios that performed better in two or more objectives. In other words, the Pareto-optimal set is the collection of scenarios in which it is not possible to improve on one objective without reducing performance in another objective. By definition, this set cannot identify a single best scenario or evaluate trade-offs between objectives; we use it here to identify obviously suboptimal scenarios.

<!-- For example, the objective function values for the scenario simulating managed aquifer recharge and in-lieu recharge (MAR and ILR, or scenario identifier `mar_ilr`) are 76.1 predicted coho smolts per female equivalent (HB value), 111.0 million m^3^ of crop ET, and infrastructure category 3. Most scenarios (33) have a higher HB value than `mar_ilr`, but only eight have a higher ET value. Only the reservoir scenarios show up in both sets, with higher ET and higher HB value than `mar_ilr`. However, the reservoir scenarios are more difficult to implement, so any of those scenarios would have less desirable performance on the infrastructure objective. Consequently, since no scenarios perform better than `mar_ilr` in more than one objective, it must be included in the Pareto-optimal set. -->

## Hydrologic model summary

The Scott Valley Integrated Hydrologic Model (SVIHM) [-@TolleyEtAlSensitivity2019] is a surface and groundwater model driven by tributary stream inflows, precipitation, and reference evapotranspiration (ET~0~). Like all models, it does not perfectly reproduce historical observations, but it successfully captures key watershed-scale hydrologic phenomena that can be used to evaluate potential outcomes of specific management actions.

<!-- The boundary of the model domain corresponds approximately to the extent of the alluvial aquifer.  -->
<!-- For detailed information about the model structure, sensitivity, calibration, uncertainty, and validation of results, see recent work by Tolley et al. .  -->

The SVIHM is composed of three cascading sub-models. First, complete daily records of tributary stream inflow are generated using a statistical streamflow regression model, in which the continuous Fort Jones record was used to predict missing values in the daily flow records of 12 major tributaries [@FogliaEtAlScott2013]. Second, a custom-built soil water budget model simulated irrigation demand, irrigation applied (surface and groundwater), ET and percolation to aquifer calculated on a field-by-field, daily time step basis. This model simulates individual irrigation applications and assumes the farmer has perfect foresight of daily irrigation demand. Finally, the boundary conditions of stream inflow, aquifer recharge, and groundwater extraction are used to drive an integrated groundwater-surface water model using MODFLOW; stream-aquifer flux and surface flow routing was simulated using the MODFLOW SFR package [@TolleyEtAlSensitivity2019]. 

For GSP development, SVIHM was used extensively to perform what-if scenarios using the historic climate and land use forcing period, covering 1991-2018. (For purposes of this study, the model period has been extended through Sept. 30, 2025.) Scenario simulations repeat the calibrated historical basecase simulation of SVIHM while varying specific changes in SVIHM boundary conditions (BCs). The set of changes in the BCs of a particular scenario represent specific changes in land use, irrigation management, climate, water management rules, etc. Scenario outcomes are compared against those of the basecase simulation. Key results of each SVIHM scenario, used in the objective functions described below, are:

1. A prediction of Scott River daily average flow at the Fort Jones Gauge, which is co-located with the surface water outlet of the model domain; and
2. Simulated total ET of crops and native vegetation within the model domain.

## Management scenario descriptions

For the Scott Valley GSP, SVIHM was used to simulate 64 management action scenarios; in most other California GSPs, between one and five scenarios was more common [e.g., FCGMA -@FoxCanyonGroundwaterManagementAgencyGroundwater2019; YSGA -@YoloSubbasinGroundwaterAgencyYolo2022]. This high number of scenarios was possible because the SVIHM had been under development for 8 years prior to the start of GSP work [e.g., @FogliaEtAlScott2013; @TolleyEtAlSensitivity2019], and cost-effective technical support was provided by graduate students at the University of California, Davis. Of the 64 scenarios, only 25 are considered here. Several were ruled out as obviously unfeasible (e.g., construction of a 134 thousand acre-foot reservoir to ensure 100% flow reliability even during severe drought, in an area with topography that could not support such a large capacity), and some were excluded because they are largely redundant with some of the 25 considered here. 

The scenarios are grouped here by category (Table \@ref(tab:scenCategoryTab)). Within each category various details were altered (e.g., an alfalfa irrigation cessation date of July 10 versus August 15). See Supplemental Material for more detailed scenario descriptions.

```{r scenCategoryTab, echo = FALSE, results = 'asis'}


#Create caption and column labels
caption_text = "Management scenario categories (abbreviated), a general description, the number of scenarios in each category, the feasibility proxy and the type of management action."
column_labels = c("Category", "Scenario Category Description", "Num. Scen.","Infra. Proxy","Mgmt. Type")

scen_cat_tab = get_scen_cat_tab()
scen_cat_tab_for_ft = scen_cat_tab #make a copy for modifying for display
colnames(scen_cat_tab_for_ft) = column_labels

scen_cat_tab_ft = flextable(data = scen_cat_tab_for_ft)
scen_cat_tab_ft = set_caption(scen_cat_tab_ft, caption = caption_text)
scen_cat_tab_ft = flextable::width(scen_cat_tab_ft,j=1,  width = 1.2)
scen_cat_tab_ft = flextable::width(scen_cat_tab_ft,j=2,  width = 1.8)
scen_cat_tab_ft = flextable::width(scen_cat_tab_ft,j=3,  width = 0.8)
scen_cat_tab_ft = flextable::align(scen_cat_tab_ft,j=3, align="center")
scen_cat_tab_ft = flextable::width(scen_cat_tab_ft,j=4,  width = 0.8)
scen_cat_tab_ft = flextable::align(scen_cat_tab_ft,j=4, align="center")
scen_cat_tab_ft = flextable::width(scen_cat_tab_ft,j=5,  width = 1.5)
scen_cat_tab_ft


```



# Results

```{r farmerFishTradeoffTableMainText, echo = F, warning =F}

scenario_tab_init = get_scenario_tab_init()
scenario_tab_init = scenario_tab_init[scenario_tab_init$scenario_category!="Attr",]
coef_tab = coef_tab[, c("Predictor", "Value")]

# Calculate mean of 35 years of hydrologic benefit value and ET
obj_fn_results = get_HBF_and_crop_ET_columns(scenario_tab = scenario_tab_init,
                                        lasso_weights = coef_tab)

scenario_tab = obj_fn_results$scenario_tab
obj_fn_wy = obj_fn_results$obj_fn_results_by_wy

# replace transformed data with straight abundance data for better interpretability

# for scenario tab
scenario_tab$hb_untrans_mean = 10^scenario_tab$HBF_mean
scenario_tab$hb_untrans_stdev = 10^scenario_tab$HBF_stdev

colnames(scenario_tab) = gsub(pattern = "HBF",
                              replacement = "HBF_log10",
                              x = colnames(scenario_tab))
colnames(scenario_tab) = gsub(pattern = "hb_untrans",
                              replacement = "HBF",
                              x = colnames(scenario_tab))

# for obj_fn_wy
obj_fn_wy$hbf_total_untrans = 10^obj_fn_wy$hbf_total
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hbf_total"] = "hbf_total_log10"
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hbf_total_untrans"] = "hbf_total"

```

As described above, objective functions were used to quantitatively summarize the achievement of environmental and agricultural goals for the basecase and `r nrow(scenario_tab) - 1` management scenarios. A Hydrologic Benefit (HB) value in units of predicted smolt abundance for coho and a total crop ET volume was calculated for each water year in the 35-year simulation period, for each scenario. (See Supplemental Material for these results calculated for an HB function with the other three units: coho smolt per adult spawner, Chinook smolt abundance, or Chinook smolt per adult.) In addition to the two objective functions, each scenario has been assigned an infrastructure feasibility category (Table \@ref(tab:scenarioResultsTabMainText)). 

<!-- ## 4.1 Environmental and agricultural objective function values -->
## Environmental and agricultural objective function values

```{r calc_num_4.1, echo = F, warning =F}

# calculate numbers for results discussion
top_2_cat = c("EnhRch", "Res")
top_2_selector = scenario_tab$scenario_category %in% top_2_cat
top_2_et_range = rev(range(scenario_tab$et_mean[top_2_selector]) / 10^6 * -1)
top_2_hbf_improve = range(scenario_tab$HBF_mean[top_2_selector] -
                            scenario_tab$HBF_mean[scenario_tab$scenario_id=="basecase"])
top_2_et_worse = -1 / 10^6 * range(
  scenario_tab$et_mean[scenario_tab$scenario_id=="basecase"] -
    scenario_tab$et_mean[top_2_selector] )

flowlims_et_worse = -1 / 10^6 *
  (scenario_tab$et_mean[scenario_tab$scenario_id=="basecase"] -
    scenario_tab$et_mean[scenario_tab$scenario_id=="eflows25_div_lims"])
flowlims_hb_better = (scenario_tab$HBF_mean[scenario_tab$scenario_id=="eflows25_div_lims"]) -
  scenario_tab$HBF_mean[scenario_tab$scenario_id=="basecase"]

alf_curts_cat = c("Curtail", "AlfIrr")
alf_curts_selector = scenario_tab$scenario_category %in% alf_curts_cat
alf_curts_et_range = rev(range(scenario_tab$et_mean[alf_curts_selector]) / 10^6 * -1)
alf_curts_hbf_improve = range(scenario_tab$HBF_mean[alf_curts_selector] -
                            scenario_tab$HBF_mean[scenario_tab$scenario_id=="basecase"])
alf_curts_et_worse = -1 / 10^6 * range(
  scenario_tab$et_mean[scenario_tab$scenario_id=="basecase"] -
    scenario_tab$et_mean[alf_curts_selector] )

bc_et_mean = scenario_tab$et_mean[scenario_tab$scenario_id == "basecase"] *-1 / 10^6
bc_hb_mean = scenario_tab$HBF_mean[scenario_tab$scenario_id == "basecase"]

scen_et_range = range(scenario_tab$et_mean[scenario_tab$scenario_id != "basecase"] *-1 / 10^6)
scen_hb_range = range(scenario_tab$HBF_mean[scenario_tab$scenario_id != "basecase"])

swbm_bc = get_swbm_tab_with_separated_et(scen_id="basecase")
swbm_nv_low = get_swbm_tab_with_separated_et(scen_id="natveg_low_all")
swbm_nv_high = get_swbm_tab_with_separated_et(scen_id="natveg_high_all")

# mean(swbm_nv$`Nat. Veg. ET`)/mean(swbm_bc$`Nat. Veg. ET`)
# mean(swbm_nvet$`Nat. Veg. ET`)/mean(swbm_nv$`Nat. Veg. ET`)


natveg_et_percent = scenario_tab$et_mean[grep(pattern = "natveg",
                                              x = scenario_tab$scenario_id)] *-1 / 10^6


```


```{r scenario_tab_plot_columnsMainText, echo = F, warning=F}

# Plot specs
scen_cat_tab = get_scen_cat_tab()
scen_cat_tab = add_plot_cols_to_scen_cat_tab(scen_cat_tab_1 = scen_cat_tab)

# Assign category featurs to full scenario table
scenario_tab$color = scen_cat_tab$color[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$feas_cat = scen_cat_tab$feas_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$mgmt_cat = scen_cat_tab$mgmt_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$symbol = scen_cat_tab$symbol[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab = add_pareto_col_to_scen_tab(scenario_tab = scenario_tab,
                                          yes_size = 15, no_size = 10)

```


```{r benefitByWYMainText, echo = F, fig.cap="\\label{fig:benefitByWY} Comparison of annual performance in two objective functions (Hydrologic Benefit and ET, representing environmental and agricultural objectives, respectively) for brood years 1991-2025. The scenarios pictured are basecase (black circles) and a scenario from each of three categories: Enhanced Recharge, Crop Change, and eight thousand acres of irrigated crops converted to low-ET Native Vegetation. Dry-type water years (vertical bars) are 1991, ’92, ’94, 2001, ’09, ’13, ’14, ’18, and ’20-’22. Shown also in Supplement A." , warning=F}

y_val_x_axis_label = model_name_to_y_val_label(model_name)

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 7.5, units = "in", res = 300)

  plot_scenarios1 = c("basecase", "maxMAR_fields2024",
                      "grain_12k", "natveg_low_8k")

  par(mfcol = c(2,1))
  plot_obj_fn_by_wy(plot_scenarios = plot_scenarios1,
                    cols = get_palette(c("black","dodgerblue",
                                         "green4","darkgoldenrod"),
                                       length(plot_scenarios1)),
                    panel_labels = c("A","B"),
                    y_val_label = y_val_x_axis_label)
  dev.off()
}

knitr::include_graphics(fig_path)
fig_i = fig_i + 1

```

```{r farmerFishTradeoffFigureMainText, echo = F, fig.cap="\\label{fig:farmerFishTradeoffFigure} Environmental and agricultural objective function performance (in terms of Hydrologic Benefit, or HB, value; and crop ET volume) for basecase (black circle in upper left) and 25 management scenarios. Symbol x and y coordinates are the average of 35 annual HB and ET values for each scenario over the full model period (water years 1991-2025); for exampe time series of annual values, see previous figure." , fig.height=8, fig.width=7, warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_fig(scenario_tab = scenario_tab,
                         y_val_label = y_val_x_axis_label,
                         plot_title = "Environmental vs Agricultural Benefit \n Mean of annual values, 1991-2025")
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1

#Make blank axes for slides
if(print_for_slides==T){png(filename = file.path(save_figs_here, paste0("Figure ",fig_i,".png")), width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_blank(scenario_tab = scenario_tab,
                           y_val_label = y_val_x_axis_label)
  fig_i = fig_i +1; dev.off()}

```

The aggregate hydrologic impact of each management scenario can be compared against the historical basecase using the annual objective function values over the full 35-year record (e.g., Figure \@ref(fig:benefitByWYMainText)) or using the 35-year mean values for each scenario (Figure \@ref(fig:farmerFishTradeoffFigure); Table \@ref(tab:scenarioResultsTabMainText)). In the basecase scenario, the 35-year mean annual ET volume is `r round(bc_et_mean)` million m^3^ of water, and the 35-year mean of annual HB values is `r round(bc_hb_mean)` coho spf-equivalent (Table \@ref(tab:scenarioResultsTabMainText)). Annual ET volume is relatively stable over the model period of water years 1991-2025, while annual HB values are much more variable (Figure \@ref(fig:benefitByWY)). Among the `r nrow(scenario_tab) - 1` management scenarios, mean ET values ranged from `r round(scen_et_range[1])` (for a no cultivated agriculture scenario) to `r round(scen_et_range[2])` million m^3^ of water, while HB values ranged from `r round(scen_hb_range[1])` to `r round(scen_hb_range[2])` coho smolt-equivalent.

**results bullet points**

* High-importance flows selected in LASSO regression and corroboration by other statistical approaches (Table \@ref(tab:coefTabMainText)). 
  + Because predictor values were standardized (Z-score transformed) before the statistical modeling exercise, absolute values of coefficients for flow metrics can be interpreted as relative degree of influence on the relevant ecological outcome.
  + For coho, in both absolute and spawner-normalized smolt units, LASSO regression identified at fall-to-wet season transition metrics as of high importance for predicting coho outcomes; for units of absolute smolt abundance, the spring recession rate was also selected (Supplements A and B, Table 1). 
  + For Chinook, in both absolute and spawner-normalized smolt units, the wet season baseflow median magnitude was of greatest importance; for normalized smolt units, maximum spring recession rate was also selected (Supplements C and D, Table 1).  
  + Effectively, in the context of LASSO regression, the flows that explain the most variation in smolt production (without overfitting, as determined by k-fold validation) were related to, for coho, the timing of wet season onset and the magnitude of changes in fall flows; and for Chinook, the wet season baseflow.
  + The importance of wet season baseflows to Chinook and fall flows to coho is corroborated by the other statistical modeling approaches, i.e., LASSO regression predicting smolt production based on hydrology and spawner abundance, as well as MARSS regression models predicting smolt production with single hydrologic covariates [@KoubaEtAlWatershedSpecific2025]. However, other metrics were also highlighted: **summarize hydro+spawner and MARSS results** [@KoubaEtAlWatershedSpecific2025].
  
* for HB over time  (Figure \@ref(fig:benefitByWYMainText) and also Supplement A, Figure 1:
  + Generally a uniform effect on ET, shifting it by about the same amount each year. Conversely, scenario effect on HB function is variable. (Supplements A-D, Figure 1)
  + For HB over time, not a lot of difference between scenarios some years. Some scenarios may not change the component flows relative to basecase (Supplements A-D, Figure 2) 

* For HB components (SUpplements A-D, Figure 2):
  + For coho juvenile abundance, in the basecase scenario, it seems that when the HB function predicts a particularly good year (**numerical definition**), it is driven by a large magnitude fall flow difference, or fall pulse. Conversely, in the years where the HB function predicts a particularly bad year, it is driven by a rapid first spring recession rate (Supplement A, Figure 2). 
  + Conversely, for coho juveniles per spawner (basecase), both good and bad years seem to be largely driven by wet season onset timing (Supplement B, Figure 2)
  + In both cases for coho, 2nd fall flow difference makes minor contributions (Supplements A and B, Figure 2), and is assigned a negative coefficient. This suggests higher flows in the second fall (for juvenile overwintering coho) are consistently associated with lower smolt production, but other factors such as first fall flow difference and wet season timing seem to exert greater overall influence.
  + For Chinook, predictions of both high and low smolt production are driven by wet season baseflow magnitude (with a small contribution from maximum spring recession rate for juveniles per adult): high wet season flows are associated with lower smolt abundance. 


* Tradeoff plot (Figure \@ref(fig:farmerFishTradeoffFigureMainText).
  + Chinook plots of fish vs farmer tradeoffs are effectively meaningless, since wet season median flows are largely beyond human control in this undammed watershed. (**for discussion?**) 
  + Highest-HB value is total conversion of irrigated agriculture to native vegetation. (High-ET vegetation endmember for Chinook and low-ET for coho).
  + reservoir scenarios don't help coho much at all: small reservoirs just don't have capacity to add fall flow during dry years when baseflows are lowest. Surprisingly, reservoir scenarios seem to help Chinook significantly more: even a small reservoir on one tributary can dampen wet season stormflows enough to benefit Chinook.
  + In all four HB formulations, Enhanced Recharge produces virtually no change in either fish benefit or farmer cost relative to basecase (Supplements A-D, Figure 3). The lack of farmer cost is by design; this management strategy is intended to minimize any impacts on growing season operations. But the lack of any fish benefit may reflect limitations of current surface water diversion permit rules; the maximum amount permitted winter diversions may not be sufficient to generate baseflow increases at the end of the following dry season.
  + For coho, Enhanced Recharge and Reservoir scenarios produce virtually no farmer cost. 
  + Irrigation efficiency seems
  
**check - is alf_curtail_07.31 actually alf_curtail_08.15?**

* Efficiency plot (Figure \@ref(fig:efficiencyMainText)
  + For coho, 
  
* 3-D results figure  (Figure \@ref(fig:farmerFishInfraFigure3D)
  + For coho, trade-off curves in every category.
 
Four scenario categories fall in a cluster with relatively high ET values, indicating that these scenarios conserve nearly all the basecase economic benefits for local growers: Enhanced Recharge, Expanded Enhanced Recharge, Irrigation Efficiency, and Small Reservoir (Figure \@ref(fig:farmerFishTradeoffFigure)). The scenarios in this group all require some degree of infrastructure investment, and they have a minimal impact on crop ET: the mean ET values of scenarios in this cluster are `r round(top_2_et_range[1])` to `r round(top_2_et_range[2])` million m^3^, or `r round(top_2_et_worse[1]/bc_et_mean*100)` percent less to `r round(top_2_et_worse[2]/bc_et_mean,3)*-1*100` percent more than basecase. Their mean improvement in HB values over the basecase ranges from `r round(top_2_hbf_improve[1])` to `r round(top_2_hbf_improve[2])` coho spf-equivalent.

The mean ET volumes of the two Crop Change scenarios fall below those in the aforementioned cluster, with `r round(scenario_tab$et_mean[ scenario_tab$scenario_id == "irrig_0.8"] / 10^6 * -1)` and `r round(scenario_tab$et_mean[ scenario_tab$scenario_id == "irrig_0.9"] / 10^6 * -1)` million m^3^, while HB values are `r round(scenario_tab$HBF_mean[ scenario_tab$scenario_id == "irrig_0.8"],1)` and `r round(scenario_tab$HBF_mean[ scenario_tab$scenario_id == "irrig_0.9"],1)` respectively (Figure \@ref(fig:farmerFishTradeoffFigure)). In these scenarios, existing crop coefficients are multiplied by a factor of 0.8 or 0.9; conceptually, this would involve replacing all existing irrigated acreage with unspecified crops which had a designated 80% and 90% of historical ET demand. As discussed later, in this scenario the relationship between water use and revenue may be distinctly different from that of alfalfa and pasture, so the actual "farmer benefit" may be higher or lower than this proxy suggests.

Two regulatory management categories, Alfalfa Irrigation and Curtailment, involve irrigation schedule changes, and result in a range of reductions in applied irrigation volume, depending on the date of intervention. The scenarios in these categories trace two trade-off arcs in average HB-ET space (Figure \@ref(fig:farmerFishTradeoffFigure)). The first category, early cessation of alfalfa irrigation (during all or only dry water years), is intended to simulate growers harvesting two cuttings only (i.e., one or two fewer cuttings than current) in a given year. This was simulated by ceasing alfalfa irrigation earlier in the growing season (with cutoff dates on July 10, August 1 and August 15) than the basecase default date of Sept. 1. (The Sept. 1 cutoff date assumes that growers obtain three cuttings, but does not incorporate the known small number of growers which occasionally irrigate longer to obtain a fourth cutting.) Alfalfa acreage is concentrated near the Scott River channel, much of it in the Adjudicated Zone (Figure \@ref(fig:scottLanduseMap)), with predominately groundwater irrigation sources. The second regulatory category, Curtailment of all irrigation starting on a range of dates (June 1st through August 15th), represent book-end scenario simulations to approximate regulatory actions proposed by state resource management agencies in response to extreme drought conditions [SWRCB -@StateWaterResourcesControlBoardEstablishment2021a]. The mean ET values of scenarios in these two categories are `r round(alf_curts_et_range[1])` to `r round(alf_curts_et_range[2])` million m^3^, or `r round(alf_curts_et_worse[1]/bc_et_mean*100)` to `r round(alf_curts_et_worse[2]/bc_et_mean,3)*100` percent less than basecase; HB gains over basecase are `r round(alf_curts_hbf_improve[1])` to `r round(alf_curts_hbf_improve[2])` coho spf-equivalent.

The third regulatory category, Low Flow Diversion Limits, regulates surface water diversion for irrigation, and contains only one scenario (Figure \@ref(fig:farmerFishTradeoffFigure)). Among all scenarios, it is the only scenario that focuses solely on management of surface water diversions for irrigation. It assumes the FJ Gauge flowrate is representative of watershed conditions, and limits diversions of surface water from all tributaries on days when the FJ Gauge flowrate falls below a recently proposed instream flow regime that is considered to be protective of anadromous fish in the Scott River [CDFW -@CaliforniaDepartmentofFishandWildlifeINTERIM2017]. Under this Low Flow Diversion Limits scenario, the mean crop ET volume is `r round(flowlims_et_worse)` million m^3^ (~`r round(flowlims_et_worse/bc_et_mean*100)`%) less than basecase, while the mean HB value is `r round(flowlims_hb_better)` coho spf-equivalent greater than basecase.

The remaining two categories, abbreviated as "NatVeg" and "NatVegET", relate to land cover type and are not proposed as management actions. Rather, these provide reference scenarios for conditions in some or all parts of the valley that could be considered to stand in for undefined "pre-historic", "unimpaired", or "conservation" conditions. They involve simulating native vegetation in place of irrigated acreage, and were designed to estimate the streamflow depletion attributable to agricultural water use in different areas of Scott Valley. These scenarios produce substantial HB gains at a high ET cost relative to basecase (Figure \@ref(fig:farmerFishTradeoffFigure)).

The key geographic feature in the NatVeg and NatVegET scenarios is the "Interconnected Zone" in the 1980 Scott Valley surface water adjudication [@CaliforniaStateWaterResourcesControlBoardScott1980], referred to here as the "adjudicated zone" of groundwater pumping (Figure \@ref(fig:scottLanduseMap)), and the six scenarios in each category represent the cessation of all irrigation or the cessation of only groundwater pumping in three spatial configurations: within the adjudicated zone, outside of the adjudicated zone, or in both areas. All areas where irrigation or pumping is turned off are assumed to return to (unirrigated) native vegetation. 

On average, area-normalized ET from native vegetation is `r round(et_nv_depth_m*100)` cm/year (`r round(et_nv_depth_m/et_crop_depth_m*100)`% of average annual crop ET).  Simulating native vegetation in place of irrigated acreage results in a greater proportion of water leaving the SVIHM model domain via surface flows rather than as transpiration. Consequently the NatVeg and NatVeg ET scenarios generate higher simulated FJ streamflow in most months of the model period. The first category (NatVeg) assumes a native vegetation crop coefficient of 0.6, while the second category (NatVegET) assumes a native vegetation crop coefficient of 1.0 and a maximum rooting depth of 4.5 m. The first category could be interpreted as chaparral or sage scrub-type vegetation cover, while the second category could represent a mixed bunchgrass and clover prairie with extensive riparian vegetation corridors and wetlands. (See *Supplemental Material* for more details on this scenario design.)


\blandscape

```{r scenarioResultsTabMainText, echo = FALSE, results = 'asis'}

# subset desired columns
scen_tab = scenario_tab[,c("scenario_id", "scenario_category", "feas_cat", 
                           "HBF_mean", "et_mean", "pareto_opt",
                           "hb_gained_to_et_lost")]
if(max(scen_tab$HBF_mean) > 2000){
  scen_tab$HBF_mean = round(scen_tab$HBF_mean / 1000, 2)
  mean_HB_text = "Mean HB / 1000"
} else {
  scen_tab$HBF_mean = round(scen_tab$HBF_mean, 2)
  mean_HB_text = "Mean HB"
}

scen_tab$et_mean = round(scen_tab$et_mean)
# remove the negative efficiencies - that means ET was gained, not lost, so it doesn't apply
scen_tab$hb_gained_to_et_lost = round(scen_tab$hb_gained_to_et_lost, 2)
scen_tab$hb_gained_to_et_lost[scenario_tab$hb_gained <=0 | 
                                scenario_tab$et_lost <=0 |
                                !is.finite(scen_tab$hb_gained_to_et_lost) |
                                scen_tab$scenario_id == "basecase" ] = "--"
scen_tab$et_mean = round(scenario_tab$et_mean * (-1 / 10^6), 2)

# Exclude attribution scenarios
scen_tab = scen_tab[scen_tab$scenario_category != "Attr",]

#Create caption and column labels
caption_text = "Scenario attributes for 25 scenarios. Values in the Mean Hydrologic Benefit (HB) and Mean ET columns are averages of 35 annual values for each scenario. Efficiency not calculated for scenarios with no average ET loss."
column_labels = c("Scenario ID", "Category", "Feas.", mean_HB_text, "Mean ET",  "Pareto", "Efficiency")

scen_tab_for_ft = scen_tab
colnames(scen_tab_for_ft) = column_labels

st_ft = flextable(scen_tab_for_ft)
st_ft = set_caption(st_ft, caption = caption_text)

st_ft = flextable::fontsize(st_ft,j=1,  size = 10)
st_ft = flextable::fontsize(st_ft,j=2,  size = 10)
st_ft = flextable::fontsize(st_ft,j=3,  size = 10)
st_ft = flextable::fontsize(st_ft,j=4,  size = 10)
st_ft = flextable::fontsize(st_ft,j=5,  size = 10)
st_ft = flextable::fontsize(st_ft,j=6,  size = 10)
st_ft = flextable::fontsize(st_ft,j=7,  size = 10)

# st_ft = flextable::rotate(st_ft, rotation="btlr",j=1:8)
st_ft = flextable::width(st_ft,j=1,  width = 1.7)
st_ft = flextable::width(st_ft,j=2,  width = 1)
st_ft = flextable::width(st_ft,j=3,  width = .7)
st_ft = flextable::width(st_ft,j=4,  width = .7)
st_ft = flextable::width(st_ft,j=5,  width = .7)
st_ft = flextable::width(st_ft,j=6,  width = .8)
st_ft = flextable::width(st_ft,j=7,  width = .8)

st_ft


``` 



<!-- ## 4.2 Agricultural-environmental trade-off efficiency -->
## Agricultural-environmental trade-off efficiency

```{r efficiency_calcs, echo = F, warning = F }

n_no_et_cost = sum(scenario_tab$et_lost < 0)

ef_min = round(min(scenario_tab$hb_gained_to_et_lost[scenario_tab$scenario_id!="basecase" &
                                                       scenario_tab$et_lost > 0]),2)
ef_max = round(max(scenario_tab$hb_gained_to_et_lost[scenario_tab$scenario_id!="basecase" & scenario_tab$et_lost > 0]),1)

# ef_min_lc = round(min(scenario_tab$hb_gained_to_et_lost_lc[scenario_tab$scenario_id!="basecase" &
#                                                        scenario_tab$et_lost > 0]),2)
# ef_max_lc = round(max(scenario_tab$hb_gained_to_et_lost_lc[scenario_tab$scenario_id!="basecase" & scenario_tab$et_lost > 0]),1)

# High efficiency scenarios
high_eff_scen = scenario_tab$scenario_id[scenario_tab$hb_gained_to_et_lost > 1 &
                                           !is.na(scenario_tab$hb_gained_to_et_lost)]
et_cost_Mm3 = bc_et_mean - (scenario_tab$et_mean[scenario_tab$scenario_id %in%
                                                   high_eff_scen]/ 10^6 * -1)
et_cost_percent = (bc_et_mean -
                     scenario_tab$et_mean[scenario_tab$scenario_id %in% high_eff_scen] *
                     -1 / 10^6  ) / bc_et_mean *100

et_cost_range_high_eff = round(range(et_cost_Mm3), 1)
et_cost_range_high_eff_percent = round(range(et_cost_percent), 1)

# low efficiency scenarios
low_eff_scen = scenario_tab$scenario_id[scenario_tab$hb_gained_to_et_lost < 1 &
                                          scenario_tab$hb_gained_to_et_lost > 0 &
                                           !is.na(scenario_tab$hb_gained_to_et_lost)]
et_cost_Mm3_low = bc_et_mean - (scenario_tab$et_mean[scenario_tab$scenario_id %in%
                                                  low_eff_scen]/ 10^6 * -1)
et_cost_percent_low = (bc_et_mean -
                     scenario_tab$et_mean[scenario_tab$scenario_id %in% low_eff_scen] *
                     -1 / 10^6  ) / bc_et_mean *100
et_cost_range_low_eff = round(range(et_cost_Mm3_low), 1)
et_cost_range_low_eff_percent = round(range(et_cost_percent_low), 1)

hb_gain_low = (scenario_tab$HBF_mean[scenario_tab$scenario_id %in% low_eff_scen] - bc_hb_mean) #/ bc_hb_mean *100
hb_gain_range_low_eff = round(range(hb_gain_low))

# overall ranges
et_cost_percent_all = (bc_et_mean - scenario_tab$et_mean * -1 / 10^6  ) / bc_et_mean *100
hb_gain_percent_all = (scenario_tab$HBF_mean - bc_hb_mean) / bc_hb_mean *100

et_cost_perc_all_range = round(range(et_cost_percent_all[et_cost_percent_all>0]),1)
hb_gain_perc_all_range = round(range(hb_gain_percent_all[et_cost_percent_all>0]),1)

# , and were higher (ranging from `r ef_min_lc` to `r ef_max_lc`) in low-coho years 
```

The trade-off efficiency between the environmental and agricultural objective functions is defined as the average HB value gained, relative to HB~basecase~, per ET value lost relative to ET~basecase~ (see formula in Methods). However, six scenarios (the five reservoir scenarios and managed aquifer recharge, or MAR) actually resulted in an average ET volume slightly greater than the basecase, and so avoided any ET cost. These scenarios were excluded from the efficiency calculations because including them would produce nonsensical negative efficiencies.

For all other scenarios, trade-off efficiencies ranged from `r ef_min` to `r ef_max` when averaged over all years(Table \@ref(tab:scenarioResultsTabMainText)). The range covers two orders of magnitude because the denominator of the ratio, lost ET, ranges from  `r et_cost_perc_all_range[1]`% to `r et_cost_perc_all_range[2]`%, while the HB gains fall more narrowly between `r hb_gain_perc_all_range[1]`% to `r hb_gain_perc_all_range[2]`%. Efficiencies greater than 1 indicate that a scenario produces more than a 1% increase in HB value per 1% loss in ET value. However, because the HB and ET units are non-commensurable, the absolute value of the trade-off efficiency is not of great significance; the trade-off efficiency metric is most useful for scenario comparisons.

Four scenario categories form a "low-efficiency" cluster, with all-years efficiencies of less than 1: Crop Change, Curtailment, and the two Natural Vegetation categories (Figure \@ref(fig:efficiency)). These scenarios can produce substantial HB gains (ranging from `r hb_gain_range_low_eff[1]` to `r hb_gain_range_low_eff[2]`) coho spf-equiv., but result in severe agricultural productivity reductions  (`r et_cost_range_low_eff_percent[1]`% to `r et_cost_range_low_eff_percent[2]`%). The "high-efficiency" scenarios, which have an all-years trade-off efficiency greater than 1 (the Enhanced Recharge, Low Flow Diversion Limits, Irrigation Efficiency and Alfalfa Irrigation scenarios), tend to be associated with a small ET cost relative to basecase (`r et_cost_range_high_eff[1]` to `r et_cost_range_high_eff[2]` million m^3^, or `r et_cost_range_high_eff_percent[1]` to `r et_cost_range_high_eff_percent[2]` percent of basecase crop ET).


```{r efficiencyMainText, echo = F, fig.cap="\\label{fig:efficiency} Scenario trade-off efficiencies, or average Hydrologic Benefit value (a proxy for environmental benefit) gained per crop ET value lost (a proxy for farmer cost) (normalized to average basecase HB and ET values, respectively), in all 35 years. Some scenarios are excluded if they have no or negative ET loss, producing infinite or negative efficiency values." , warning=F}

bc_sel = scenario_tab$scenario_id=="basecase"
hb_bc_mean = scenario_tab$HBF_mean[bc_sel]
et_bc_mean = scenario_tab$et_mean[bc_sel]
scenario_tab$hb_gained = (scenario_tab$HBF_mean - hb_bc_mean)/hb_bc_mean
scenario_tab$et_lost = (scenario_tab$et_mean - et_bc_mean)/et_bc_mean *-1

scenario_tab$hb_gained_to_et_lost = round(scenario_tab$hb_gained / scenario_tab$et_lost, 2)



fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 6, units = "in", res = 300)
  tradeoff_efficiency_fig_nov2025(scenario_tab = scenario_tab)
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1




```

<!-- ## 4.3 Infrastructure objective, Pareto-optimal set and GSP priorities -->
## Infrastructure objective, Pareto-optimal set and GSP priorities

By plotting the environmental, agricultural, and infrastructure objectives in a three-dimensional scatterplot, it is possible to visualize the combined set of management scenarios as four different trade-off arcs in different infrastructure categories (Figure \@ref(fig:farmerFishInfraFigure3D)). 

The Pareto-optimal set of management actions include more than half (25) of the simulated scenarios (Table \@ref(tab:scenarioResultsTabMainText); Figure \@ref(fig:farmerFishInfraFigure3D)). All the suboptimal scenarios are in infrastructure category 2 (infrastructure timeline of 1-2 years). The suboptimal set are either related to land use change (Crop Change, NatVeg, and NatVegET), which reduce crop ET demand or irrigated acreage, or they involve deployment of more efficient irrigation (Irrigation Efficiency), which has little impact on either HB or ET values. 

Infrastructure category 1 (no infrastructure investment needed) contains all the Regulation-type scenarios. In the Alfalfa Irrigation and Curtailment scenario categories, the trade-off arcs show the effect of intervention date: an earlier date of intervention (irrigation cutoff) produces a greater HB value at a higher ET cost, compared to a later date of intervention. The scenarios in infrastructure categories 3 and 4 (infrastructure timeline of 3-5 and 5+ years) involve no or minimal ET cost, and a range of HB gains on the middle-to-lower end of the HB gain spectrum (Reservoirs, Enhanced Recharge, and Expanded Enhanced Recharge; Figure \@ref(fig:farmerFishTradeoffFigure)). 

To determine if the strict application of objective functions would produce different management priorities than the public policy development process that took place 2018-2021, we made a high-level, conceptual comparison of the Pareto-optimal set of management actions and the management actions and projects listed for consideration in the GSP [Siskiyou County -@SiskiyouCountyScott2021]. Management actions considered in the GSP have some analogs in the set of 25 simulated scenarios. 

High-priority scenarios tend to be associated with higher average crop ET values and modest HB gains relative to basecase (Table \@ref(tab:gspMgmtList)); i.e., the GSP priorities demonstrate a preference for management actions with high trade-off efficiencies. Specifically, Enhanced Recharge and Irrigation Efficiency are represented in Tiers I and II, signifying they are considered implementable in the short- or medium-term. Another high-efficiency scenario, Low Flow Diversion Limits, is represented three times in Table \@ref(tab:gspMgmtList): the two analogs for it in Tier I are limited either in time (Water Trust Leasing, which occurs opportunistically in extreme droughts) or in space (the current two-creek extent of the existing watermaster program). An expanded watermaster program, which would more closely resemble the Low Flow Diversion Limits scenario, appears in Tier III. 

The final high-efficiency scenario category, Alfalfa Irrigation, is not included in the list of management actions; conversely, low-efficiency scenarios in categories NatVeg and Crop Change were included, in a spatially-limited form, in Tier II under the umbrella of "Voluntary Managed Land Repurposing". This indicates that political considerations not currently included in the three objective functions produce a preference for some amount of land repurposing rather than imposing limitations on alfalfa irrigation.

Notably, direct comparisons between the set of simulated scenarios and the management actions listed in the GSP are somewhat limited. The primary reason for this is that the simulated scenarios assume that all water users in the valley are acting in concert. Conversely, management actions included in the GSP generally assumed voluntary/incentivized participation by individual landowners in scenarios involving land use change, and in some cases are located only in specific sub-areas of the watershed.


```{r farmerFishInfraFigure3D, echo = F, fig.cap="\\label{fig:farmerFishInfraFigure3D} A three-dimensional scatter plot of the objective function values for basecase and 39 management scenarios, averaged over 28 water years (1991-2018). Larger spheres are shown for the Pareto-optimal set of scenarios. Scenario colors correspond to legend in Figures 6 and 7." , out.width="80%", warning=F}


save_3d_plot = F

if(save_3d_plot == T){
  # need to run pareto_set_calc R chunk first
plot3d(x = scenario_tab$HBF_mean, 
       y = scenario_tab$et_mean / (-1 * 10^6), 
       z = scenario_tab$feas_cat,
       type = "s", col = scenario_tab$color, 
       radius = scenario_tab$size_3d,
       xlab = "HB",#"Hyd. Ben Value",
       ylab = "ET vol.",
       zlab = "Infrastructure")
}
#This only works to manually orient it and take a screen shot; do this if you need to redo this graph

scatter3d_filename = file.path(fig_dir, "hbf et infra plot3d_2026.02.19.jpg")
knitr::include_graphics(scatter3d_filename)

```


```{r gspMgmtList, echo = FALSE, results = 'asis'}

gsp_tab = read.csv(file.path(data_dir,"scott_gsp_mgmt_action_list_2025.11.04.csv"))
gsp_tab$hbv_range = NA; gsp_tab$et_range = NA

for(i in 1:nrow(gsp_tab)){
  scens = trim(unlist(strsplit(gsp_tab$Scenario_id[i], split = ",")))
  scen_selector = scenario_tab$scenario_id %in% scens
  if(length(scens)<2){
      gsp_tab$hbv_range[i] = round(scenario_tab$HBF_mean[scen_selector],2)
      gsp_tab$et_range[i] = round(scenario_tab$et_mean[scen_selector] * -1 / 10^6)
  } else {
    gsp_tab$hbv_range[i] = paste(round(range(scenario_tab$HBF_mean[scen_selector]),2), collapse = "-")
    gsp_tab$et_range[i] = paste(round(range(scenario_tab$et_mean[scen_selector] * -1 / 10^6)), collapse = "-")
  }
}

gsp_tab = gsp_tab[,c("Tier", "Title", #"Description",
                     "Similar.Scenario.Categor.y.ies.", "hbv_range", "et_range")]

#Create caption and column labels
caption_text = "List of management actions in Chapter 4 of the GSP and analogues in the set of 25 simulated scenarios. The Tier indicate the priority given to each management action, though the Tier also incorporates information about difficulty of implementation (hence Reservoirs being placed in Tier 3)."
column_labels = c("Tier", "GSP Management Action", #"Description",
                  "Closest Simulated Analogue", "HB Range", "ET Range")
gsp_for_ft = gsp_tab
colnames(gsp_for_ft)=column_labels

#declare xtable object, assign formatted column labels, and print to latex object
# gsp_xtable = xtable(gsp_tab, label = "tab:gspMgmtList",
#                  caption = caption_text, #align = "llp{2.7cm}p{6cm}p{2.7cm}ll"
#                  #align = "llp{4cm}p{8cm}p{4cm}ll"
#                  align = "llp{5cm}p{4.3cm}ll")#, digits = c(NA, NA, NA, NA, NA, NA, NA))
# names(gsp_xtable) = column_labels
# print.xtable(gsp_xtable, comment = F,type = "latex", format.args = list(big.mark = ","), include.rownames = F, size="\\fontsize{10pt}{12pt}\\selectfont", hline.after = c(0, 3, 7, 12))
#caption.placement = "top",

gsp_ft = flextable(gsp_for_ft)
gsp_ft = set_caption(gsp_ft, caption = caption_text)
gsp_ft = flextable::width(gsp_ft,j=1, width = .5)
gsp_ft = flextable::width(gsp_ft,j=2, width = 2)
gsp_ft = flextable::width(gsp_ft,j=3, width = 2.2)
gsp_ft = flextable::width(gsp_ft,j=4, width = .9)
gsp_ft = flextable::width(gsp_ft,j=5, width = .9)
gsp_ft

``` 

\newpage

# Discussion

**add discussion of alternative env obj functions: steelhead, other aquatic species, GDEs, indices of alteration**

**add discussion of how to interpret grain ET results. might need to check with TH**

The portfolio of scenarios proposed by stakeholders and simulated for this study included a range of infrastructure investments, regulatory actions and major land use changes. The number and diversity of simulated scenarios was made possible by a surface- and groundwater model designed to approximate existing land use and agricultural water use behavior in the local context [SVIHM; @TolleyEtAlSensitivity2019]. Though the ultimate future effect of these management actions would depend on climate variability and implementation details that have not been explicitly simulated (e.g., the precise location of pumps needed to divert surface water for the expanded enhanced recharge scenarios), it is possible to use objective functions and historical climate records to draw some high-level findings regarding their relative merits.

## Contextualizing low objective function values

The HB function incorporates several hydrologic metrics, but years with low HB values (i.e., low predicted coho-spf equivalent) usually share several characteristics. 

A water year with a low HB value (e.g., < XX coho spf-equiv.) commonly occurs after a dry season with multiple dried-out mainstem Scott River reaches, culminating in a prolonged fall period (e.g., 4 to 8 weeks after September 1st) of FJ Gauge flows lower than 10 cfs. In a low-HB water year, high winter baseflows (e.g., > 200 cfs) may not arrive until December or January, and the rainy season  may be short (e.g., < 100 days). These hydrologic features have implications for fishery stability. In water year 2014, for example, salmon access flows were significantly delayed, and coho salmon laid their eggs in the mainstem of the Scott River rather than in their preferred tributary stream habitat, which put them at higher risk of getting scoured away in a winter storm [CDFW -@CaliforniaDepartmentofFishandWildlifeEtAlCooperative2015]. 

Because no recent precedent exists for a low-crop-ET year, interpreting the numerical values of the ET objective function requires more speculation, and the specific impact of ET loss on the agricultural community depends on the specific management action and its enforcement. One method to address uncertainty related to impact of management actions on agricultural objectives is to survey regional agricultural practitioners and experts about the likely costs of a proposed management action [@SundingMeasuring1996]. In the public meetings during GSP development, some informal (i.e., non-quantitative) surveying of stakeholder expectations of the costs of management actions took place. In terms of operational impacts, Scott Valley agricultural stakeholders stated that  short-term interventions (such as fallowing an alfalfa field for one year, or curtailing pasture irrigation for a growing season) could have long-term impacts (such as affecting alfalfa yield the following year, or forcing a rancher to sell a herd [SVGAC -@SiskiyouCountyGSA-ScottValleyGroundwaterAdvisoryCommitteeSiskiyou2021]. The current agricultural benefit proxy, annual crop ET, does not capture these multi-year costs to the local agricultural sector, or the regional impact that reduced agricultural productivity will have on economically disadvantaged local communities. 

These predicted on-the-ground consequences can facilitate interpretation of scenario results.

## Environmental objective performance in historically low-coho years

```{r calc_num_5.2, echo = F, warning =F}

# calculate numbers for results discussion
# ET reduction from curtailment scenario
aet_bc = aet_by_lu_budgets[["basecase"]]#get_aet_by_landuse_table(scenario_id = "basecase")
aet_bc$Month = seq.Date(from = start_date, by = "month", length.out=nrow(aet_bc))
aet_bc$Water_Year = wtr_yr(aet_bc$Month)
et_crop_bc = aet_bc$Alfalfa_Irrigated + aet_bc$Grain_Irrigated + aet_bc$Pasture_Irrigated

aet_curtail07.31 = aet_by_lu_budgets[["curtail_all_07.31"]]#get_aet_by_landuse_table(scenario_id = "curtail_start_07.31")
aet_curtail07.31$Month = seq.Date(from = start_date, by = "month", 
                                  length.out=nrow(aet_curtail07.31))
aet_curtail07.31$Water_Year = wtr_yr(aet_bc$Month)

et_crop_curtail07.31 = aet_curtail07.31$Alfalfa_Irrigated + 
  aet_curtail07.31$Grain_Irrigated +
  aet_curtail07.31$Pasture_Irrigated

# aggregate to annual
aet_bc_annual = mean(aggregate(et_crop_bc,
                               by = list(aet_bc$Water_Year),
                               FUN = sum)$x)
aet_ca01_annual = mean(aggregate(et_crop_curtail07.31,
                            by = list(aet_curtail07.31$Water_Year),
                            FUN = sum)$x)
crop_et_reduce_curtail_07.31 =  (aet_bc_annual - aet_ca01_annual)/aet_bc_annual

# HB improvement from curtailment scenario
# scenario selectors
bc_selector =  obj_fn_wy$scenario=="basecase"
scen1_selector = obj_fn_wy$scenario=="curtail_all_07.31"
scen2_selector = obj_fn_wy$scenario=="eflows25_div_lims"
scen3_selector = obj_fn_wy$scenario == "maxMAR_fields2024"
scen4_selector = obj_fn_wy$scenario=="alf_curtail_08.15"
scen5_selector = obj_fn_wy$scenario=="alf_curtail_07.31"
scen6_selector = obj_fn_wy$scenario=="alf_curtail_07.15"
# years selector
# low_hb_selector = obj_fn_wy$wy %in% low_bc_hb_yrs
# 
# bc_hb_lc = obj_fn_wy$hbf_total[bc_selector & low_hb_selector]
# scen1_hb = obj_fn_wy$hbf_total[scen1_selector & low_hb_selector]
# scen2_hb = obj_fn_wy$hbf_total[scen2_selector & low_hb_selector]
# scen3_hb = obj_fn_wy$hbf_total[scen3_selector & low_hb_selector]
# scen4_hb = obj_fn_wy$hbf_total[scen4_selector & low_hb_selector]
# scen5_hb = obj_fn_wy$hbf_total[scen5_selector & low_hb_selector]
# scen6_hb = obj_fn_wy$hbf_total[scen6_selector & low_hb_selector]
# 
bc_et = obj_fn_wy$et_tot[bc_selector]
scen1_et = obj_fn_wy$et_tot[scen1_selector]
scen2_et = obj_fn_wy$et_tot[scen2_selector]
scen3_et = obj_fn_wy$et_tot[scen3_selector]
scen4_et = obj_fn_wy$et_tot[scen4_selector]
scen5_et = obj_fn_wy$et_tot[scen5_selector]
scen6_et = obj_fn_wy$et_tot[scen6_selector]

# numbers for discussion
# range(scen1_hb - bc_hb_lc)
# aug1_hb_improve_lc = mean(scen1_hb - bc_hb_lc)
# # range(scen2_hb - bc_hb_lc)
# flowlim_hb_improve_lc = mean(scen2_hb - bc_hb_lc)
# mar_hb_improve_lc = mean(scen3_hb - bc_hb_lc)
# alf_08.15_hb_improve_lc = mean(scen4_hb - bc_hb_lc)
# alf_07.31_hb_improve_lc = mean(scen5_hb - bc_hb_lc)
# alf_07.15_hb_improve_lc = mean(scen6_hb - bc_hb_lc)

aug1_et_worse = mean(scen1_et - bc_et) / 10^6
flowlim_et_worse = mean(scen2_et - bc_et) / 10^6
# crop_et_reduce_flowlim = (mean() - aet_ca01_annual)/aet_bc_annual
mar_et_worse = mean(scen3_et - bc_et) / 10^6
alf_08.15_et_worse = mean(scen4_et - bc_et) / 10^6
alf_07.31_et_worse = mean(scen5_et - bc_et) / 10^6
alf_07.15_et_worse = mean(scen6_et - bc_et) / 10^6


# Among Alfalfa Irrigation scenarios, HB gains in low-coho years are modest under cutoff dates of August 15th and August 1st (averaging `r round(alf_08.15_hb_improve_lc)` and `r round(alf_07.31_hb_improve_lc)` more coho spf than basecase in low-coho years, respectively),

 # and improves HB performance only slightly (`r round(mar_hb_improve_lc)` coho spf gain over basecase in low-coho years)
```


Management scenario performance in dry years, rather than over the whole model period, is another informative metric for resource managers. For example, in potentially unstable fish populations, conditions during critical dry years can be more significant than long-term average conditions  [e.g., @OhlbergerEtAlEffects2018]. Eight dry years occurred during the 1991-2018 model period: 1991, '92, '94, 2001, '09, '13, '14, and '18 (Figure \@ref(fig:benefitByWY)). Dry years were classified using quartiles of the distribution of total annual FJ Gauge flow over the full record of water years 1942-2021. The lowest to highest quartiles of total annual flow are classified as Dry, Below Average, Above Average, and Wet water year types, respectively.

However, in this system, the years following dry periods may be more informative. As mentioned in Methods, the lowest basecase HB values, operationally defined here as years with basecase HB values of less than XX predicted coho spf-equivalent, tend to occur in years following one or more dry years (i.e., 1993, 1995, 2002, and 2014-16 in Figure \@ref(fig:benefitByWY); "low-coho years"). In the Scott River watershed, one mechanism producing this pattern is delayed onset of winter storms, which cannot be managed through policy. Another mechanism is the depletion of stored groundwater in dry years, which can reduce dry-season baseflow and delay stream system reconnection during the onset of the following wet season [@TolleyEtAlSensitivity2019]. Regardless of the primary cause of the low fall flows in a particular year, the HB value for each water year is sensitive to long periods in the fall with flow below 10 and 100 cfs (see Chapter 1 of this dissertation for more details on coho salmon sensitivity to low fall flows).

Because many aspects of scenario performance are relatively consistent (i.e., for any given scenario, ET costs are relatively static year to year, and most scenarios demonstrate consistently small HB gains over basecase in high-coho spf years; Figure \@ref(fig:benefitByWY)), the low-coho year HB performance of various model scenarios is a key scenario differentiator.

## Environmental-agricultural trade-offs

Balancing agricultural and environmental water demands in this watershed commonly takes the form of seeking management actions that can improve flow conditions while minimizing the resulting reduction in agricultural water availability. Consequently, a subset of the proposed scenarios rely on infrastructure investments to produce modest HB improvements with minimal-to-no ET reductions (Enhanced Recharge, Expanded Enhanced Recharge, Irrigation Efficiency, and Reservoirs) (Figure \@ref(fig:farmerFishTradeoffFigure)). The remaining categories tend to include some level of ET reduction, due to reduced water availability for irrigation.

The environmental gain-agricultural cost trade-off is evident both within individual categories and between distinct scenario categories, especially in HB performance in low-coho years.

Among Alfalfa Irrigation scenarios,  and ET costs are small (averaging `r round(alf_08.15_et_worse)` and `r round(alf_07.31_et_worse)` million m^3^, respectively, in all years); conversely, the more extreme July 10th cutoff date produces an average of `r round(alf_07.15_et_worse)` million m^3^, or `r round(alf_07.15_et_worse/(et_bc_mean *-1 / 10^6)*100)`%, reduction in ET in all years (panels A and B of Figure \@ref(fig:benefitByWY)).

Trade-off is also evident between scenario categories. The Enhanced Recharge scenario `mar_ilr` has a minimal ET cost (an average of `r round(mar_et_worse,1)` million m^3^ less than basecase in all years). At the other extreme, the Curtailment scenario in which all irrigation is cut off on August 1st (`curtail_start_07.31`), but reduces ET by an average of `r round(aug1_et_worse)` million m^3^ or `r round(crop_et_reduce_curtail_07.31*100)`% (panels C and D of Figure \@ref(fig:benefitByWY)). Compared to `curtail_start_07.31`, the Low Flow Diversion Limits (`flowlims`) scenario reduces ET by an average of `r round(flowlim_et_worse / (bc_et_mean)*100)`%.


The 33 scenarios with ET lower than basecase sort clearly into one high- and one low-efficiency cluster (Figure \@ref(fig:efficiency)). The high-efficiency scenarios are composed of three infrastructure-type management actions (Enhanced Recharge, Expanded Enhanced Recharge and Irrigation Efficiency) and two regulatory-type management actions (Alfalfa Irrigation and Low Flow Diversion Limits), while the low-efficiency group contains one infrastructure-type (Crop Change), one regulatory-type (Curtailment), and the two native vegetation-type scenarios. The previously-stated caveat regarding the Crop Change scenario's true agricultural benefit potentially being higher or lower than the values in Table \@ref(tab:scenarioResultsTabMainText) apply to these efficiency calculations as well. The trade-off efficiencies provide an informal cost-benefit assessment, though interpretation is somewhat limited, as the unitless quantity tends to group together scenarios with widely varying absolute values of HB and ET (e.g., Crop Change and some Curtailment scenarios).

There is a clear positive relationship between efficiency in all years and in low-coho years, but in nearly all scenarios the low-coho year efficiency average is higher than over the 35-year model period (Figure \@ref(fig:efficiency)). In other words, in low-coho years, a given amount of crop ET reduction produces a greater environmental flow enhancement than on average over all years. This is corroborated by annual scenario HB values: in high-coho years, HB gains are small or minimal, while substantial improvements are seen for some scenarios in low-coho years (panels A and C in Figure \@ref(fig:benefitByWY)). One possible explanation is that, in high-coho years, environmental flows tend to already be sufficient, and most management scenarios can generate only marginal improvements in hydrologic conditions. This suggests that targeting management actions to dry years (which tend to precede the low-coho water years) may produce the highest efficiency management actions. This possibility was supported by the two cases where this comparison was explicitly simulated: the efficiencies for the Alfalfa Irrigation scenarios for dry years only (`alf_irr_stop_07.31_dry_yrs_only` and `alf_irr_stop_08.15_dry_yrs_only`) are higher than for their corresponding scenarios with an intervention in all years (`alf_irr_stop_07.31` and `alf_irr_stop_08.15`; Table \@ref(tab:scenarioResultsTabMainText)).

## Pareto-optimal solution set and GSP priority management actions

Within the proposed suite of management actions, more than half (**25** of 25, including the basecase) fall within the Pareto-optimal set based on the infrastructure category and the 35-year average values of objective functions. This same set remains Pareto-optimal when the objective functions are averaged only over the six low-coho water years, with one exception (the French Creek reservoir scenario is Pareto-optimal when averaged in all years, but not in low-coho years).

The non-Pareto-optimal scenarios (i.e. the scenarios outperformed in multiple objectives) are in the Irrigation Efficiency, Crop Change and Native Vegetation categories. (The 20% improved Irrigation Efficiency scenario is outperformed in both HB and ET objectives by both of the alfalfa irrigation scenarios with a cutoff date of August 15; Table \@ref(tab:scenarioResultsTabMainText)). Native Vegetation scenarios, under two different assumptions about vegetation type/ET demand of the vegetation replacing cultivated agriculture, can produce significant HB improvements, but at a relatively high cost in regional agricultural productivity; these two scenario categories have the lowest trade-off efficiency (Figure \@ref(fig:efficiency); Table \@ref(tab:scenarioResultsTabMainText)).

<!-- (Figure \@ref(fig:scottLanduseMap)) -->

Pareto-suboptimal status is less certain in the case of the Crop Change scenario, since the ET volume proxy for agricultural revenue/benefit may not be valid for alternative crops. (Crops that have been mentioned during stakeholder meetings as possibly suited to the climate of Scott Valley or have been cultivated in a trial setting are high(er) value annuals in lieu of the current perennial crops, e.g., potatoes, sunflowers, and carrots for seed, though notably, to the authors’ knowledge no local grower has cultivated any of these commercially.)

Many of the simulated scenarios were included in the ultimate list of management priorities in the adopted GSP [Siskiyou County -@SiskiyouCountyScott2021], though in a variety of forms (e.g., the Low Flow Diversion Limits scenario, which is the closest simulated analogue for a watermaster program and a surface water leasing program; Table \@ref(tab:gspMgmtList)). Although the GSP management priorities emphasized infrastructure, in a strict application of these objective functions, both infrastructure-based and regulatory approaches fell within the Pareto-optimal set of management options.

## Policy implications

Although the HB values are based on well-studied hydro-biological relationships (late spawning flows leads to lower fishery reproduction rates), the actual predicted values of coho smolt production per female spawner (coho spf) remain uncertain, because when the HB function was developed only 11 full years of coho spf data were available. If monitoring continues in future years and this sample size increases, the predictive power and weights of various hydrologic metrics may be reexamined. The ET values are estimated based on climate data, crop coefficients, and assumptions regarding grower decisions, and thus are subject to uncertainty in those data sources.

Due to this uncertainty, the results of this study should not be considered direct policy recommendations. However, some of the insights below could be used to support future management decisions, simply by summarizing the wealth of information provided in each model simulation using the proposed objective functions.

* Targeted interventions that only regulate water use in dry periods produce higher trade-off efficiencies than continuous interventions, as illustrated by the Alfalfa Irrigation scenarios (i.e., dry years only versus all years; Table \@ref(tab:scenarioResultsTabMainText)). 
* The Low Flow Diversion Limits scenario takes this concept to its logical extreme, by limiting intervention only to low-flow days in the hydrologic record, rather than low-flow years or seasons in the Alfalfa Irrigation category. As a result it produces considerable low-coho year HB gains with a trade-off efficiency of greater than 1.
* Key data gaps limit interpretability of some scenario results. An agronomic study of alternate crops that could be cultivated in Scott Valley, for example, could better constrain the agricultural benefits of a Crop Change scenario. Conversely, the acreage of land converted in various Native Vegetation scenarios is likely higher than any realistic land use change scenario, meaning that for a hypothetical management action involving a reduction of irrigated acreage in the valley, the HB gains and ET costs may be smaller than shown in Figure \@ref(fig:farmerFishTradeoffFigure). 
* In recent years, drought conditions have prompted state regulatory agencies to consider and then impose significant irrigation curtailments [SWRCB -@StateWaterResourcesControlBoardEstablishment2021a; -@StateWaterResourcesControlBoardEstablishment2024a; -@STATEWATERRESOURCESCONTROLBOARDRESOLUTION2024]. The objective function summary of existing model results could help stakeholders and regulators weigh the trade-offs of various actions and potential irrigation cutoff dates in the context of the consequences of other proposed management scenarios. The modeling results suggest that if a decision is made to regulate water use, some later intervention dates may produce minimal gains in environmental flows in some water years (e.g., an alfalfa irrigation cutoff date of August 15 in water years 1995 or 2002 in panel A of Figure \@ref(fig:benefitByWY)). 


# Conclusion

In this study, functions quantifying the achievement of the two management objectives (environmental and agricultural) were applied to 25 simulated management scenarios (including a historical basecase). In addition, scenarios were assigned a categorical proxy for new infrastructure costs. We used these numerical values to quantitatively explore trade-offs in environmental and agricultural benefit among the suite of proposed actions.

Due to decades of investment and significant ongoing labor, aggregate agricultural productivity in Scott Valley (as estimated through crop ET) is buffered against the weather and is relatively stable year-to-year. Conversely, flow in the undammed Scott River is subject to the full range of climate variability and human water uses. As a result, the historical performance in the environmental flows objective is much more variable than in the agricultural objective, and environmental flows measured with this objective function seem particularly vulnerable to multi-year dry periods.

Many management actions were proposed over the past four years to enhance environmental flows while minimizing negative effects on agricultural water security. In general, targeting irrigation reductions during low-flow periods (e.g. Early Cessation of Alfalfa Irrigation and Low Flow Diversion Limits) tends to be more efficient in terms of higher environmental flows per lost agricultural productivity than irrigation reduction over the whole growing season, such as the Native Vegetation scenarios (Figure \@ref(fig:efficiency)).

Objective functions summarizing the predicted consequences of management actions is not a new technique; however, to be useful in a decision support context, they require careful design, sensitivity to local conditions, and ongoing feedback from affected stakeholders. Introducing these elements of optimization into existing water resource management processes can improve communication of management consequences and priorities, and thus potentially improve decision-making and our understanding of the interconnected socio-hydrologic system. 




\newpage

# Acknowledgments {-}
<!-- The authors are grateful for feedback from two reviewers and financial support from the California State Water Resources Control Board Agreement, #19-12-110 and #21-066-300; and the Nora S. Gustavsson Endowed Professorship for Groundwater Resources in Agriculture, University of California Davis. -->

# References

<div id="refs"></div>

\newpage

# Supplement: Fish-Farm Results from all four Hydrologic Benefit function formulations


For each species, across all statistical methods, the coefficients for any single predictor were consistently negative or positive; e.g., 

The Hydrologic Benefit (HB) function is intended to estimate the ecological value (in units of fish production) provided to fish by specific flow values (e.g., timing of fall flow increase; magnitude of dry season flow). This is expressed in conceptual mathematical terms as $fish = b + m_1 * flow_1 + m_1*flow_2~{\cdots}$; see methods in  @KoubaEtAlWatershedSpecific2025 for more details. It's important to note that, in this dataset, only 40-60% of variation in fish outcome was explained in models based on flow data, and the HB function does not constitute a prediction of the number of fish outmigrating in a given year. 

Specifically, the results below are based on statistical models, using LASSO regression, of reproductive output for two species of salmon (coho and Chinook) in the Scott River stream system. Reproductive output is expressed in units of outmigrating juvenile abundance and normalized juveniles-per-adult spawner; two reproductive units for two species yields the four HB function formulations. 

While MARSS models were used as a tool for identifying important functional flows in [@KoubaEtAlWatershedSpecific2025], they are unsuitable for this hydrologic scenario outcome forecasting purpose, because they are autoregressive in nature: the forecasting of future fish outcomes requires historical observations of fish outcomes. Therefore, to use MARSS models in estimating the ecological effects of flow changes under different management scenarios, it would be necessary to have historical fish records collected in an alternate reality under the different management scenarios, which is not possible. Similarly, LASSO models which included spawner abundance as a predictor were also not included in this exercise. 

Consequently, only LASSO models dependent on hydrologic variables were evaluated as potential tools to forecast the outcomes of flow change on fish reproduction.

<!-- ## Coho, juvenile abundance -->

<!-- ```{r varDeclaration1, echo=F, include=F} -->
<!-- model_name = "coho_lasso_juv_on_hydro" -->
<!-- ``` -->

<!-- ```{r call_supplement1, child=if (INCLUDE_SUPP) 'Kouba_Fish_and_Ag_Benefit_supplement.Rmd'} -->
<!-- ``` -->

<!-- ## Coho, juveniles per spawner -->

<!-- ```{r varDeclaration2, echo=F, include=F} -->
<!-- model_name = "coho_lasso_jps_on_hydro" -->

<!-- ``` -->

<!-- ```{r call_supplement2, child=if (INCLUDE_SUPP) 'Kouba_Fish_and_Ag_Benefit_supplement.Rmd'} -->
<!-- ``` -->

<!-- ## Chinook, juvenile abundance -->

<!-- ```{r varDeclaration3, echo=F, include=F} -->
<!-- model_name = "chinook_lasso_juv_on_hydro" -->

<!-- ``` -->

<!-- ```{r call_supplement3, child=if (INCLUDE_SUPP) 'Kouba_Fish_and_Ag_Benefit_supplement.Rmd'} -->
<!-- ``` -->

<!-- ## Chinook, juveniles per spawner -->

<!-- ```{r varDeclaration4, echo=F, include=F} -->
<!-- model_name = "chinook_lasso_jps_on_hydro" -->

<!-- ``` -->

<!-- ```{r call_supplement4, child=if (INCLUDE_SUPP) 'Kouba_Fish_and_Ag_Benefit_supplement.Rmd'} -->
<!-- ``` -->
