---
title: "Chapter 2 - Scenario Description Supplement"
author: "Claire Kouba"
date: "3/29/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  toc: yes
header-includes:
  - \usepackage{pdflscape}
  - \usepackage{rotating}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
fig_caption: yes
bibliography: library.bib

---

```{r setup_dirs_load_data, include = FALSE}

  # - \usepackage{setspace}\doublespacing
  # - \usepackage[left]{lineno}
  # - \linenumbers

knitr::opts_chunk$set(echo = TRUE)


library(lubridate)
library(rgdal)
library(rpostgis)
library(postGIStools)
library(here)
library(httr)
library(stringr)
library(rstudioapi)
library(dataRetrieval) # for usgs data
library(data.table) #month function
library(zoo) # rolling means
library(plyr)  # for barplots for climate fig
library(rgl) # plot3d
library(ggplot2) # water budget bar chart
#spatial data libraries
library(rgeos)
library(maptools)
library(RANN)
library(tmap)
library(RColorBrewer)
library(raster)
library(xtable)
library(scales) # for alpha() color function
library(colorspace) # for sequential_hcl()
library(ggpubr) # to use get_palette()

# Directories
# diss_dir <- dirname(getActiveDocumentContext()$path ) 
diss_dir = here::here()
scratch_dir = file.path(diss_dir, "scratch_work")
# Directory for postprocessed, tabulated results
svihm_results_dir = file.path(diss_dir, "Data", "SVIHM Model Results")
# Directories for SVIHM inputs and outputs, not yet postprocessed
svihm_scenarios_dir = "C:/Users/Claire/Documents/GitHub/SVIHM/Scenarios"
svihm_scenarios_backup_dir = "G:/SVIHM Scenarios Storage"

# Load spatial and tabular data
# If the data files don't exist locally,
if(!file.exists(file.path(diss_dir,"dissertation_data.RData"))){
  source(file.path(diss_dir,"01_Database connection.R"))   # connect to AWS server
  source(file.path(diss_dir,"02_DataRetrieval_Cleaning_SaveLocal.R")) #pulls data off server, local files, and internet
}
load(file.path(diss_dir, "dissertation_data.RData")) # Load data layers/tables
source(file.path(diss_dir,"03_Ch1_Figure_Functions.R"))   # Load figure functions
source(file.path(diss_dir,"04_Ch2_Figure_Functions.R"))   # Load figure functions

# print figures for slides, yes or no?
save_figs_here = "C:/Users/Claire/Dropbox/Documents/UCD/Presentations_Talks_Workshops_miniprojects/2022.02.17 Peer Group presentation Ch. 1/figs"
print_for_slides=F; fig_i = 1

```

<!-- # to do -->

<!-- * Additionally, each description inncludes a justification regarding its assigned cost/feasibility category. -->



# Introduction to Scenario Descriptions

To support development of the Scott Valley Groundwater Sustainability Plan [GSP; Siskiyou County -@SiskiyouCounty2021], a team of technical specialists, including the authors of this study, generated the following 40 management action scenarios. The scenarios are grouped here by Management Type (one of four options: Basecase, Infrastructure, Regulatory, and Native Vegetation) and by Scenario Category (\autoref{tab:scen_category_tab}). A full list of scenario identifiers is shown in \autoref{tab:scenario_results_tab}. 

Within each category various details were altered (e.g., an alfalfa irrigation cessation date of July 10 versus August 15, or an assumed crop water consumption of 80% versus 90% of historical). The scenario descriptions below will follow this format:

# *X. Management Type*

## *X.X Scenario Category* ***(Abbrev.)***

*[Scenario category description and rationale]*

* **scenario_identifier** -- *[Specific scenario details]*




```{r scen_category_tab, echo = FALSE, results = 'asis'}


#Create caption and column labels
caption_text = "Management scenario categories (abbreviated), a general description, the number of scenarios in each category, the feasibility proxy and the type of management action."
column_labels = c("Category", "Scenario Category Description", "Num. Scen.","Feas. Proxy","Mgmt. Type")

scen_cat_tab = get_scen_cat_tab()

#declare xtable object, assign formatted column labels, and print to latex object
scen_cat_xtab = xtable(scen_cat_tab, label = "tab:scen_category_tab",
                 caption = caption_text, digits = c(NA, NA, NA,0,0,NA) )#,align = "lp{6cm}p{6cm}"))
names(scen_cat_xtab) = column_labels
print.xtable(scen_cat_xtab, comment = F,type = "latex", format.args = list(big.mark = ","), include.rownames = F)#, hline.after = 0:3)
#caption.placement = "top",

```

```{r initialize_scenario_tab, echo = F, warning =F}


scenario_tab_init = get_scenario_tab_init()

# Calculate weights based on best models

# Read in slopes table for all predictors
coef_tab = read.csv(file.path(diss_dir, "Graphics and Supplements",
                              "best linear model coefficients table_2022 feb.csv"))
colnames(coef_tab) = c("model_id", colnames(coef_tab)[2:(ncol(coef_tab))])
# Averaged row for HBF weights
metric_weights = t(coef_tab[coef_tab$model_id=="Avg.", 2:(ncol(coef_tab))])

# Calculate mean of 28 years of hydrologic benefit value and ET 
obj_fn_results = calculate_HBF_and_crop_ET(scenario_tab = scenario_tab_init,
                                        weights = metric_weights,
                                        include_years = "all")

scenario_tab = obj_fn_results$scenario_tab
obj_fn_wy = obj_fn_results$obj_fn_results_by_wy
```

```{r scenario_tab_plot_columns, echo = F, warning=F}

# Plot specs
scen_cat_tab = add_plot_cols_to_scen_cat_tab(scen_cat_tab_1 = scen_cat_tab)

# Assign category featurs to full scenario table
scenario_tab$color = scen_cat_tab$color[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$feas_cat = scen_cat_tab$feas_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$mgmt_cat = scen_cat_tab$mgmt_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$symbol = scen_cat_tab$symbol[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]

```

```{r initialize_and_plot_cols_for_scenario_tab_dry, echo = F, include = F, warning=F}

# calculate dry water years (by quartiles over 1942-2021)
fj_wy_type = fj_wy_type_tab(fjd = fj_flow)
dry_years = fj_wy_type$wy[fj_wy_type$type == "Dry"]

# Calculate mean of 28 years of hydrologic benefit value and ET
dry_wy_results = calculate_HBF_and_crop_ET(scenario_tab = scenario_tab_init,
                                        weights = metric_weights,
                                        include_years = dry_years)

scenario_tab_dry = dry_wy_results$scenario_tab
scenario_tab_dry$color = scen_cat_tab$color[match(scenario_tab_dry$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab_dry$feas_cat = scen_cat_tab$feas_cat[match(scenario_tab_dry$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab_dry$symbol = scen_cat_tab$symbol[match(scenario_tab_dry$scenario_category,
                                                    scen_cat_tab$category)]


```


```{r pareto_set_calc, echo = F, warning=F}

scenario_tab = add_pareto_col_to_scen_tab(scenario_tab = scenario_tab)
scenario_tab_dry = add_pareto_col_to_scen_tab(scenario_tab = scenario_tab_dry)


```


\blandscape

```{r scenario_results_tab, echo = FALSE, results = 'asis'}
# attach dry year numbers to scenario table


# Calculate efficiency - HB gained to ET lost
scenario_tab$hb_et_ratio = round(scenario_tab$HBF_mean / scenario_tab$et_mean * 10^6 * -1, 1)
scenario_tab$et_hb_ratio = round(scenario_tab$et_mean / 10^6 * -1 / scenario_tab$HBF_mean, 1)
bc_sel = scenario_tab$scenario_id=="basecase"
scenario_tab$hb_gained = scenario_tab$HBF_mean - scenario_tab$HBF_mean[bc_sel]
scenario_tab$et_lost = (scenario_tab$et_mean - scenario_tab$et_mean[bc_sel]) / 10^6 *-1*-1
# scenario_tab$et_lost[scenario_tab$et_lost < 0] = 0
scenario_tab$hb_gained_to_et_lost = round(scenario_tab$hb_gained / scenario_tab$et_lost, 2)

# Calculate efficiency - HB gained per ET lost for dry years
scenario_tab_dry$hb_et_ratio = round(scenario_tab_dry$HBF_mean / scenario_tab_dry$et_mean * 10^6 * -1, 1)
scenario_tab_dry$et_hb_ratio = round(scenario_tab_dry$et_mean / 10^6 * -1 / scenario_tab_dry$HBF_mean, 1)
bc_sel_dry = scenario_tab_dry$scenario_id=="basecase"
scenario_tab_dry$hb_gained = scenario_tab_dry$HBF_mean - scenario_tab_dry$HBF_mean[bc_sel]
scenario_tab_dry$et_lost = (scenario_tab_dry$et_mean - scenario_tab_dry$et_mean[bc_sel]) / 10^6 *-1*-1
# scenario_tab_dry$et_lost[scenario_tab_dry$et_lost < 0] = 0
scenario_tab_dry$hb_gained_to_et_lost = round(scenario_tab_dry$hb_gained / scenario_tab_dry$et_lost, 2)

# Add dry year columns to scenario tab
scenario_tab$HBF_mean_dry = scenario_tab_dry$HBF_mean
scenario_tab$et_mean_dry = scenario_tab_dry$et_mean
scenario_tab$pareto_opt_dry = scenario_tab_dry$pareto_opt
scenario_tab$hb_gained_to_et_lost_dry = scenario_tab_dry$hb_gained_to_et_lost

# subset desired columns
scen_tab = scenario_tab[,c("scenario_id", "scenario_category", "feas_cat", 
                           "HBF_mean", "et_mean", "pareto_opt", "hb_gained_to_et_lost",
                           "hb_gained_to_et_lost_dry")]
# remove the negative efficiencies - that means ET was gained, not lost, so it doesn't apply
scen_tab$hb_gained_to_et_lost = round(scen_tab$hb_gained_to_et_lost, 1)
scen_tab$hb_gained_to_et_lost_dry = round(scen_tab$hb_gained_to_et_lost_dry, 1)
scen_tab$hb_gained_to_et_lost[scen_tab$hb_gained_to_et_lost<0 | 
                                scen_tab$scenario_id == "basecase" ] = "--"
scen_tab$hb_gained_to_et_lost_dry[scen_tab$hb_gained_to_et_lost_dry<0 |
                                    scen_tab$scenario_id == "basecase"] = "--"
scen_tab$et_mean = round(scenario_tab$et_mean * (-1 / 10^6), 2)
# scen_tab$et_mean_dry = round(scenario_tab$et_mean_dry * (-1 / 10^6), 2)

#Create caption and column labels
caption_text = "Scenario attributes for 40 scenarios. Units for Hydrologic Benefit Value (HBV) are coho smolt per female-equivalent; for ET are million cubic meters; both are averages of 28 annual values. Efficiency not calculated for scenarios with no average ET loss. (Dry) indicates the efficiency is averaged over the 8 dry years that occurred within that period, rather than the full model period 1991-2018."
column_labels = c("Scenario ID", "Category", "Feas.", "Mean HBV", "Mean ET",  "Pareto", "Efficiency", "Eff. (Dry)")

#declare xtable object, assign formatted column labels, and print to latex object
scen_xtable = xtable(scen_tab, label = "tab:scenario_results_tab",
                 caption = caption_text, digits = c(NA, NA, NA, 0, 1, 1, NA, 1, 1))
names(scen_xtable) = column_labels
print.xtable(scen_xtable, comment = F,type = "latex", format.args = list(big.mark = ","), include.rownames = F, size="\\fontsize{9pt}{10pt}\\selectfont")
#caption.placement = "top",

``` 

\elandscape

# 1. Infrastructure scenarios

## 1.1 Enhanced Recharge (EnhRch) and Expanded Enhanced Recharge (EnhRchEx)

The enhanced recharge scenarios were developed to estimate the dry season streamflow increases that could be engineered by banking surface water in the aquifer, or using surface water to delay the use of groundwater pumping. 

Managed Aquifer Recharge (MAR) involves the diversion of winter surface flow for purposes of applying it to a land area and allowing it to inflitrate into the aquifer. MAR diversions are designed to take place December-March, when surface flow is generally abundant, on a set of fields selected using two criteria: a potentially-willing landowner (as of the scenario design period, 2013-2017) and ease of water application (red and orange areas, \autoref{fig:mar_ilr_map}, Panel A). Each MAR field was assigned an infiltration rate based on the SSURGO soil properties database (Tolley et al., *in prep*).

In Lieu Recharge (ILR) involves using surface water for irrigation in lieu of groundwater, starting in the early growing season (March-April) and for as long as surface water is available. This practice both delays extracting groundwater and contributes some additional recharge to the aquifer as infiltration through the crop rooting zone. Fields on which this kind of irrigation flexibility is possible by definition have access to both surface water and groundwater irrigation sources (yellow and orange areas, \autoref{fig:mar_ilr_map}, Panel A). Some fields were considered candidates for both enhanced recharge techniques (orange areas, \autoref{fig:mar_ilr_map}, Panel A). 

The Expanded Enhanced Recharge scenario design assumed that, with additional infrastructure, any field that had access to a surface water irrigation source could be included in a combined MAR and ILR program (\autoref{fig:mar_ilr_map}, Panel B). However, the additional fields in the expanded set were not assigned an infiltration rate in the same way. Instead, a range of possible average infiltration rates was tested in three scenarios: 3, 19, and 35 cm/day. 

An infrastructure category of 3 was assigned to the Enhanced Recharge and Expanded Enhanced Recharge scenarios, reflecting the authors' professional judgement that the infrastructure investments necessary for full build-out of the proposed management action would take three to five years.

```{r mar_ilr_map, echo = F, warning =F, fig.cap="\\label{fig:mar_ilr_map} Fields selected for implementation of MAR and ILR enhanced recharge practices (after Tolley et al., in prep) (Panel A), and the expanded version assuming all fields with a surface water source participate in the MAR and ILR program (Panel B)." , fig.height=8, fig.width=7}


# Plot
if(print_for_slides==T){png(file.path(save_figs_here, "Basecase Annual ET over time.png"), 
                            width = 6, height = 5, units = "in", res = 300)}

mar_ilr_map_fig()

if(print_for_slides==T){dev.off()}
```

* **mar** -- Divert surface water to over-irrigate fields and enhance groundwater recharge during the wet season (Dec-Mar). Allow diversions from tributaries to continue as long as water is available (on a monthly volume basis).

* **ilr** --  Divert surface water to irrigate fields during the growing season (Apr-Jun or Jul) in lieu of pumping groundwater. Allow diversions from tributaries to continue as long as water is available (on a monthly volume basis).

* **mar_ilr** -- Combination of MAR and ILR scenarios.

* **mar_ilr_max_0.003** -- MAR and ILR on all old MAR and ILR fields and all other fields with a surface water irrigation source. Assumed max infiltration rate of 0.003 m/day for all fields without specified infiltration rates.

* **mar_ilr_max_0.019** -- MAR and ILR on all old MAR and ILR fields and all other fields with a surface water irrigation source. Assumed max infiltration rate of 0.019 m/day for all fields without specified infiltration rates.

* **mar_ilr_max_0.035** -- MAR and ILR on all old MAR and ILR fields and all other fields with a surface water irrigation source. Assumed max infiltration rate of 0.035 m/day for all fields without specified infiltration rates.



## 1.2 Crop Change (decreased irrigation demand; CropCh)

The Crop Change scenarios were developed to assess hydrologic impacts of replacing all existing irrigated crops with an unspecified different crop that consumed 90% and 80% of the historical crop cover. To simulate this, a 0.8 or 0.9 multiplier was applied to the daily crop ET value for each field in the model domain.

An infrastructure category of 2 was assigned to the Crop Change scenarios, reflecting the authors' professional judgement that the new crops could be planted in one to two years, though many factors could change the time necessary to make this transition (for example, if the new crops are perennial rather than annuals). Crops that have been mentioned during stakeholder meetings as possibly suited to the climate of Scott Valley or have been cultivated in a trial setting include potatoes, sunflowers, and carrots for seed, though notably, to the authors’ knowledge no local grower has cultivated any of these commercially.

* **irrig_0.8** -- Assumes unspecified irrigated crop change, reducing all irrigated acreage water demand by 20%.
* **irrig_0.9** -- Assumes unspecified irrigated crop change, reducing all irrigated acreage water demand by 10%.


## 1.3 Irrigation Efficiency (IrrEff)

The Irrigation Efficiency scenarios were developed to test the hydrologic impact of widespread deployment of higher-efficiency irrigation technologies. The exact technology is unspecified; in this scenario design, a multiplier of 1.1 or 1.2 was applied to the irrigation efficiency parameter associated with each field in the model domain. 

Existing irrigation techniques in Scott Valley include flood irrigation, sub-irrigation (in areas with known shallow groundwater), wheel line, and center pivot [@Harter2008]. In recent years some agricultural operations have installed higher-efficiency technology like Mid- or Low-Elevation Spray Application, and Low Energy Precision Application has been tested in some areas [@Galdi2021].

An infrastructure category of 2 was assigned to the Irrigation Efficiency scenarios, reflecting the authors' professional judgement that, if all parties were incentivized and voluntarily participated in such a program, individual growers could install new irrigation technologies in one to two years. However, grower interest and willingness to participate in such a program is unlikely to be universal. 

* **irr_eff_improve_0.1** -- Effective irrigation efficiency of wheel line and center pivot on alfalfa and pasture improves by 0.1 (10%).
* **irr_eff_improve_0.2** -- Effective irrigation efficiency of wheel line and center pivot on alfalfa and pasture improves by 0.2 (20%).


## 1.4 Reservoirs (Res)

The Reservoir scenarios were developed to assess whether a small reservoir, located on one of the main tributary streams, could significantly enhance dry season flows. Four tributary locations were tested (from north to south): Shackleford, Etna, and French Creeks, and the South Fork of the Scott River (\autoref{fig:reservoir_map}). In each scenario the daily flow of one tributary is altered to simulate storage and release of flow. This alteration occurs upstream of the SVIHM model domain, such that from the perspective of the SVIHM, the only change from basecase is to replace one tributary's inflow to the model [i.e., the daily tributary flow estimated in the Streamflow Regression Model; @Foglia2013a; @Tolley2019] with the outflow record of the reservoir.

In four of the scenarios, the capacity of the reservoir was set at 9 thousand acre-feet (TAF, or `r round(9000 * af_to_m3 / 10^6,1)` million m^3^). This was calculated by multiplying a desired outflow, 30 cubic feet per second (cfs; `r round(30 * cfs_to_m3sec, 2)` cms or `r round(30 * cfs_to_m3day / 1000)` thousand m^3^ per day), for 150 days of a potential dry season (August-December). It also accounted for stakeholder input (as discussed during GSP development) that in the Scott Valley topographic context, a reservoir of greater than 10 TAF would probably be infeasible. However, in some multi-year drought periods, the 9 TAF reservoir runs dry and cannot meet the desired dry season outflow. To extend the analysis, an additional scenario was designed to furnish a 30 cfs dry season release with 100% reliability for the climate record 1991-2018, by ignoring the reservoir capacity constraints. The minimum reservoir capacity that accomplished this was a 29 TAF reservoir. It was sited on Etna creek to allow comparison with an existing 9 TAF reservoir scenario.

For the Reservoir scenarios, a series of operating rules dictates the following: 

1. Hold all water except 30 cfs back in the wet season (Dec. 1-Mar. 31), until the reservoir is full.
2. Allow water to pass through during the growing season (Apr. 1-June 31), to make sure enough surface water was available for irrigation, but retains water in storage.
3. Release 30 cfs in the dry season (July 1-Nov. 30), unless the reservoir runs dry.

Additionally, no feasibility or risk factors were evaluated, some of which could completely preclude actual reservoir construction. For example, stakeholder feedback regarding the Etna Creek reservoir indicated that any reservoir upstream of the community of Etna would create an unacceptable potential flood risk. Furthermore some stakeholders conveyed strenuous objections to on-stream storage solutions, due to unacceptable impacts to aquatic habitat, though off-stream storage options generated less opposition.

An infrastructure category of 4 was assigned to the Reservoir scenarios, reflecting the authors' professional judgement that the construction of any significant water storage infrastructure would take more than 5 years.


```{r reservoir_map, echo = F, warning =F, fig.cap="\\label{fig:reservoir_map} Approximate theoretical location of simulated reservoirs." , fig.height=8, fig.width=7}


# Plot
if(print_for_slides==T){png(file.path(save_figs_here, "Basecase Annual ET over time.png"), 
                            width = 6, height = 5, units = "in", res = 300)}

reservoir_map_fig()

if(print_for_slides==T){dev.off()}
```


* **reservoir_shackleford** -- Simulates a 9 TAF reservoir on the Shackleford Creek tributary by withholding wet-season flow and releasing it during the dry season according to set operations rules.
* **reservoir_etna** -- Simulates a 9 TAF reservoir on the Etna Creek tributary by withholding wet-season flow and releasing it during the dry season according to set operations rules.
* **reservoir_french** -- Simulates a 9 TAF reservoir on the French Creek tributary by withholding wet-season flow and releasing it during the dry season according to set operations rules. 
* **reservoir_sfork** -- Simulates a 9 TAF reservoir on the South Fork tributary by withholding wet-season flow and releasing it during the dry season according to set operations rules.
* **reservoir_etna_29kAF** -- Simulates a 29 TAF reservoir on the Etna Creek tributary by withholding wet-season flow and releasing it during the dry season according to set operations rules.


# 2. Regulatory scenarios

## 2.1 Early cessation of alfalfa irrigation (AlfIrr)

The Alfalfa Irrigation scenarios were developed to assess the potential for increased dry season flows in the event of a reduced alfalfa harvest (\autoref{fig:alf_irr_bounce}).

The basecase default cessation date for alfalfa irrigation in the SVIHM is August 31st. Alfalfa is a perennial crop, but is harvested periodically during "cuttings." The most common schedule in Scott Valley involves three cuttings per growing season. One scenario explored during GSP development was a proposition that alfalfa growers reduce the number of cuttings to two and then ceasing to irrigate the alfalfa crop; this would still generate some revenue for the grower but would avoid irrigation and crop ET during the critical part of the dry season. Depending on the year and growing conditions, in the two-cutting scenario, the date of the final irrigation may range from as early as mid-July to mid-August, so SVIHM was used to simulate a range of three cessation dates. A variation on this tested the impact of implementing early cessation in dry years only, which in the model period were designated as water years 1991, '92, '94, 2001, '09, '13, '14, '18. In this scenario it is assumed that a dry year can be identified before the onset of the growing season (e.g. depending on wet season precipitation and snowpack accumulation.)

An infrastructure category of 1 was assigned to the Alfalfa Irrigation scenarios, as no additional infrastructure would be required to implement this management action.

* **alf_irr_stop_jul10** -- Alfalfa irrigation ceases on July 10th of every growing season.
* **alf_irr_stop_aug01** -- Alfalfa irrigation ceases on August 1st of every growing season.
* **alf_irr_stop_aug15** -- Alfalfa irrigation ceases on August 15th of every growing season.
* **alf_irr_stop_aug01_dry_yrs_only** -- Alfalfa irrigation ceases on August 1st in designated dry years only.
* **alf_irr_stop_aug15_dry_yrs_only** -- Alfalfa irrigation ceases on August 15th in designated dry years only.

```{r alf_irr_bounce, echo = F, warning =F, fig.cap="\\label{fig:alf_irr_bounce} Effect of alfalfa irrigation scenarios on Fort Jones flow in an example water year. Because 2015 was not a dry-type year, no differences from basecase flow are seen in the two dry-year-only scenarios." , fig.height=8, fig.width=7}

# Plot
if(print_for_slides==T){png(file.path(save_figs_here, "Basecase Annual ET over time.png"), 
                            width = 6, height = 5, units = "in", res = 300)}


alf_scens = c("alf_irr_stop_aug15", "alf_irr_stop_aug15_dry_yrs_only",
               "alf_irr_stop_aug01", "alf_irr_stop_aug01_dry_yrs_only",
               "alf_irr_stop_jul10")
                
alf_colors = rev(get_palette(c("darkorchid","dodgerblue",
                               "green4","darkgoldenrod1","orangered"),
                                     length(alf_scens)))

irr_regulation_fig(scenario_ids = alf_scens, 
                   scenario_colors = alf_colors,
                   water_year = 2015,
                   plot_title = "Alfalfa Irrigation Scenario Flow Comparison")

```


## 2.2 Curtailment of all irrigation mid-growing season (Curtail)

The Curtailment scenarios were developed to assess the potential for increased dry season flows in the event of a total irrigation cut-off, sometime during the growing season  (\autoref{fig:curtail_bounce}).

The basecase default cessation date for alfalfa irrigation in the SVIHM is August 31st, while grain is irrigated through late July and pasture irrigation continues through October. Emergency drought orders by the California State Water Resources Control Board (SWRCB) in water year and 2022 [SWRCB -@SWRCB2021a; SWRCB -@SWRCB2021b] prompted a stepwise analysis, by date of initiation of curtailment, of the impact of total irrigation curtailment. 

An infrastructure category of 1 was assigned to the Curtailment scenarios, as no additional infrastructure would be required to implement this management action.

In the six Curtailment scenarios, all water years, all irrigation proceeds normal until the following curtailment dates, when all irrigation ceases:

* **curtail_start_jun01** -- June 1st
* **curtail_start_jun15** -- June 15th
* **curtail_start_jul01** -- July 1st
* **curtail_start_jul15** -- July 15th
* **curtail_start_aug01** -- August 1st
* **curtail_start_aug15** -- August 15th

```{r curtail_bounce, echo = F, warning =F, fig.cap="\\label{fig:curtail_bounce} Effect of irrigation curtailment scenarios on Fort Jones flow in an example water year." , fig.height=8, fig.width=7}

curt_scens = c("curtail_start_jun01", "curtail_start_jun15",
               "curtail_start_jul01", "curtail_start_jul15",
               "curtail_start_aug01", "curtail_start_aug15")
curt_colors = get_palette(c("violet","blue3","dodgerblue",
                               "chartreuse2","chartreuse4","darkgoldenrod3"),
                                     length(curt_scens))
  
irr_regulation_fig(scenario_ids = curt_scens, 
                   scenario_colors = curt_colors,
                   water_year = 2015,
                   plot_title = "Curtailment Scenario Flow Comparison")

if(print_for_slides==T){dev.off()}

```

## 2.3 Limits on diversion of surface water during low-flow periods at the Fort Jones Gauge (FlowLims)

The Low Flow Diversion Limits scenario was simulated to assess the hydrologic impact of restricting surface water diversions during periods when the Fort Jones Gauge falls below a proscribed desired flow regime (\autoref{fig:flowlims_hydrograph}). 

In this scenario, available water is defined as the proportion of total flow at the FJ gauge in excess of CDFW 2017 recommended instream flow values [CDFW -@CDFW2017]; alternative flow regimes could be substituted in future analyses. The “available” percentage is applied to the flow in each tributary and used to limit surface flow diversions. Surface water rights are not accounted for in this scenario. It was included in GSP development to explore the outcome of management actions such as surface water leases or potential expanded implementation of the existing water master programs on some tributary streams.

An infrastructure category of 1 was assigned to the Low Flow Diversion Limits scenario, as no additional infrastructure would be required to implement this management action.


* **flow_lims** -- Surface water diversions limited based on FJ flow and CDFW Recommended Instream Flows [CDFW -@CDFW2017]. 


```{r flowlims_hydrograph, echo = F, warning =F, fig.cap="\\label{fig:flowlims_hydrograph} Days on which, in the Low Flow Diversion Limits scenario, tributary surface water diversions for irrigation would be allowed (green x) and not allowed (red x) in an example water year." , fig.height=6, fig.width=7}

fj_streamflow_by_wy_figure(show_instream_flow_violations = T, show_simulated_flow = F)

```


# 3. Native Vegetation scenarios

The Native Vegetation scenarios were simulated to estimate the answer the following question: in 1991-2018, how much streamflow depletion was due to groundwater pumping in the area under SGMA jurisdiction? 

The area under SGMA jurisdiction is the area within the Scott River Valley Groundwater Basin that falls *outside* the area of regulated groundwater extraction, as illustrated in the Scott River adjudication decree [@SuperiorCourtofSiskiyouCounty1980] (referred to as the "Interconnected Zone" in the decree, and herein referred to as the "Adjudicated Zone") (\autoref{fig:adj_zone_map}).

Because attributing any one hydrologic phenomenon to one specific water use is difficult in such an interconnected hydrologic system, the authors of this study designed a what-if scenario to approximate conditions when groundwater pumping outside the Adjudicated Zone was eliminated during the model period 1991-2018. (Because field  boundaries do not match the Adjudicated Zone boundary, this analysis assumes that all fields with 5% or more of their area overlapping with the Adjudicated Zone are "inside" the Adjudicated Zone.) It was assumed that, on fields currently irrigated by groundwater or by combined groundwater and surface water sources, the complete cessation of groundwater pumping would make agriculture nonviable; consequently those fields would revert to unirrigated, native vegetation types. This scenario design was used to generate a new hypothetical Fort Jones Gauge flowrate in this water use case, and the attributable streamflow depletion was then calculated as the flow difference between the lower-pumping scenario and the basecase.

After this streamflow depletion attribution was calculated, some interest was expressed during stakeholder meetings in attributing streamflow depletion from to several other water uses: 1) all irrigation (i.e. not just groundwater pumping) in the area outside the Adjudicated Zone, 2) groundwater pumping and all irrigation inside the Adjudicated Zone, and 3) groundwater pumping and all irrigation on all cultivated acres in Scott Valley. Because effects of these water uses can be compounding and non-additive, this necessitated a total of six scenarios, relating the cessation of pumping and all irrigation water uses with all three configurations of land relative to the Adjudicated Zone (\autoref{tab:nv_scenarios_tab}; \autoref{fig:nv_scenarios_fig}).

All of these Native Vegetation scenarios are extremely uncertain, largely because the type of vegetation replacing crops is unknown and thus has uncertain water demand parameters (i.e., the crop coefficient k~c~ and maximum rooting depth D~ext~; further explanation below). The precise water fluxes that would take place on formerly cultivated acreage if water use ceased immediately are highly dependent on how long the vegetation has to establish, microclimate conditions, and other ecological factors. To account for this uncertainty a second set of six Native Vegetation scenarios was simulated assuming higher-ET vegetation types (category NatVegET).

An infrastructure category of 2 was assigned to the Native Vegetation scenarios, reflecting the authors' professional judgement that the conversion to native vegetation would occur in one to two years.


```{r adj_zone_map, echo = F, warning =F, fig.cap="\\label{fig:adj_zone_map} Land cover type and relevant zones used to design Native Vegetation scenarios." , fig.height=8, fig.width=7}


# Plot
if(print_for_slides==T){png(file.path(save_figs_here, "Basecase Annual ET over time.png"), 
                            width = 6, height = 5, units = "in", res = 300)}

adj_zone_map = function(){
  plot(svihm_fields, col = svihm_fields$landuse_color, border = NA,
       main = "Basecase land use and boundaries \n of Adjudicated Zone and Discharge Zone")
  plot(adj_zone, border = color_adj, add=T, lwd =2)
  plot(discharge_zone, border = color_dz, add=T, lwd = 2.5)

  descrips = lu_df$descrip
  descrips[descrips=="ET_noIrr"] = "ET, no Irr. (Nat. Veg.)"
    descrips[descrips=="noET_noIrr"] = "No ET, no Irr. (pavement, \ncobble fields)"
  
  # add discharge zone
  legend(x= "bottomleft", pt.bg = c(lu_df$color,NA, NA), 
         pch = c(rep(22,nrow(lu_df)),NA, NA), 
         pt.cex =  c(rep(2,nrow(lu_df)),NA, NA),
         legend = c(descrips, "Adjudicated Zone", "Discharge Zone"),
         lwd = c(rep(NA,nrow(lu_df)), 2, 2.5),
         lty = c(rep(NA,nrow(lu_df)), 1, 1),
         col = c(rep(NA,nrow(lu_df)), color_adj, color_dz),
         box.col=NA)

}

adj_zone_map()
if(print_for_slides==T){dev.off()}

```

```{r nv_scenarios_tab, echo = FALSE, results = 'asis'}

#Create caption and column labels
caption_text = "Explanation of various Native Vegetation scenario features. The intention regarding Attribution is to estimate the amount of streamflow depletion that would be reversed if the irrigation change under Attribution was implemented, i.e., to attribute streamflow reductions to geographically specific irrigation practices."

column_labels = c("Scenario ID", "Land Cover Description", "Flow Change Attribution", "Acres Converted")


nv_scen_tab = data.frame(scen_id =  c("natveg outside adj","natveg gwmixed outside adj",
                                      "natveg inside adj","natveg gwmixed inside adj",
                                      "natveg all","natveg gwmixed all"),
                         lu_desc = c("Nat. Veg on all fields outside Adjudicated Zone",
                                     "Nat. Veg on all fields outside Adjudicated Zone which have a groundwater or mixed groundwater-surface water irrigation source",
                                     "Nat. Veg on all fields inside Adjudicated Zone",
                                     "Nat. Veg on all fields inside Adjudicated Zone which have a groundwater or mixed groundwater-surface water irrigation source",
                                     "Nat. Veg on all irrigated fields in Scott Valley",
                                     "Nat. Veg on all irrigated fields ins Scott Valley which have a groundwater or mixed groundwater-surface water irrigation source"),
                         interp = c("Cease all irrigation outside Adjudicated Zone",
                                     "Cease all pumping outside Adjudicated Zone",
                                    "Cease all irrigation inside Adjudicated Zone",
                                     "Cease all pumping inside Adjudicated Zone",
                                    "Cease all irrigation in Scott Valley",
                                     "Cease all pumping outside in Scott Valley"),
                         acres = c(23070, 11630, 10980, 9900, 34040, 21530))


#declare xtable object, assign formatted column labels, and print to latex object
nv_scen_xtab = xtable(nv_scen_tab, label = "tab:nv_scenarios_tab",
                 caption = caption_text, digits = c(NA, NA, NA,NA,0),
                 align = "lp{2.5cm}p{5cm}p{5cm}p{1.7cm}")
names(nv_scen_xtab) = column_labels
print.xtable(nv_scen_xtab, comment = F,type = "latex", 
             format.args = list(big.mark = ","), include.rownames = F, 
             hline.after = 0:6)
#caption.placement = "top",


```


```{r nv_scenarios_fig, echo = F, warning =F, fig.cap="\\label{fig:nv_scenarios_fig} Land uses under various configurations in the Native Vegetation scenario category. Land use colors are the same as the previous figure." , fig.height=8.9, fig.width=7}

if(print_for_slides==T){png(file.path(save_figs_here, "Basecase Annual ET over time.png"), 
                            width = 6, height = 5, units = "in", res = 300)}

nat_veg_scenario_map = function(scenario_id = "natveg_all", 
                                long_title = "Native vegetation on all cultivated acres"){
  
  # Initialize column for plotting land use
  svihm_fields$landuse_nv = svihm_fields$landuse
  
  # Note: Does not convert urban land (4) or water surfaces (6) to native vegetation (3).
  # Converts only alfalfa and pasture (25, 2)
  # Water source codes: SW = 1; GW = 2; GW + SW (Mixed) = 3
  # 4 = sub-irrigated; 5 = Dry; 999 = unknown
  if(scenario_id == "natveg_outside_adj"){
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2) &
                              svihm_fields$Field_ID %in% fields_outside_adj$Polynmbr] = 3
  } else if(scenario_id == "natveg_gwmixed_outside_adj"){
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2) &
                              svihm_fields$Field_ID %in% fields_outside_adj$Polynmbr &
                              svihm_fields$wat_source %in% c(2, 3)] = 3
  }else if(scenario_id=="natveg_inside_adj"){
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2) &
                              svihm_fields$Field_ID %in% fields_inside_adj$Polynmbr] = 3
  } else if(scenario_id=="natveg_gwmixed_inside_adj"){
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2)  &
                              svihm_fields$Field_ID %in% fields_inside_adj$Polynmbr &
                              svihm_fields$wat_source %in% c(2, 3)] = 3
  } else if(scenario_id=="natveg_all") {
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2)] = 3 
  } else if(scenario_id=="natveg_gwmixed_all"){
    svihm_fields$landuse_nv[svihm_fields$landuse %in% c(25, 2)  &
                              svihm_fields$wat_source %in% c(2, 3)] = 3
  }
  
  svihm_fields$landuse_color_nv = lu_df$color[match(svihm_fields$landuse_nv, lu_df$lu_code)]
  plot(svihm_fields, col = svihm_fields$landuse_color_nv, border = NA)
  lu_extent = extent(svihm_fields)
  xmod = (lu_extent@xmax - lu_extent@xmin) * .4
  raster::text(x = lu_extent@xmin - xmod, y = lu_extent@ymin, 
               labels = long_title, srt = 90, pos = 4, cex=1.2)
    # legend(x= "bottomleft", pt.bg = c(lu_df$color,NA), 
    #      pch = c(rep(22,nrow(lu_df)),NA), pt.cex =  c(rep(1.5,nrow(lu_df)),NA),
    #      legend = c(lu_df$descrip, "Adjudicated Zone"),
    #      lwd = c(rep(NA,nrow(lu_df)), 2), col = c(rep(NA,nrow(lu_df)), color_adj))

}


nv_scen_ids = c("natveg_outside_adj","natveg_gwmixed_outside_adj",
                "natveg_inside_adj","natveg_gwmixed_inside_adj",
                "natveg_all","natveg_gwmixed_all")
titles = c("A. Native Vegetation on all fields \n outside Adjudicated Zone",
           "B. Native Vegetation on fields \n outside Adjudicated Zone \n irrigated with GW and GW+SW",
           "C. Native Vegetation on all fields \n inside Adjudicated Zone",
           "D. Native Vegetation on fields \n inside Adjudicated Zone \n irrigated with GW and GW+SW",
           "E. Native Vegetation on all \n cultivated acreage",
           "F. Native Vegetation on all \n cultivated acreage \n irrigated with GW and GW+SW")

par(mfrow = c(3,2))
for(i in 1:6){
  par(mar = c(0,0,0,0))
  nat_veg_scenario_map(scenario_id = nv_scen_ids[i], 
                       long_title = titles[i])
}


if(print_for_slides==T){dev.off()}
```



## 3.1 Native vegetation replaces agricultural acreage, various configurations (NatVeg)

Daily actual ET is explicitly estimated in SVIHM, by multiplying the daily reference ET (ET~0~) value and crop coefficients for each of the three main crops (alfalfa, pasture, grain) and native vegetation. In the basecase scenario these four vegetation covers are assigned k~c~ values of 0.95, 0.90, 0--1.15 (depending on the point in the grain growth cycle), and 0.6, respectively [@Foglia2013a; @Tolley2019].

In most of the model domain, it is assumed that crops and native vegetation draw water only from the soil zone. However, in one area, the groundwater is quite shallow and groundwater sometimes discharges to the surface. This is referred to as the "Discharge Zone," and much of the basecase land cover in the area is subirrigated pasture (\autoref{adj_zone_map}). To account for plant roots tapping this shallow groundwater, the ETS package in MODFLOW [@Banta2000] is used to simulate ET directly from the aquifer in this zone. A key parameter in the ETS package is the "extinction depth" D~ext~, analogous to the maximum rooting depth; this parameter reflects the fact that ET volume decreases with depth and is "extinguished" at the maximum vegetation rooting depth. Because the seasonal shallow groundwater in the Discharge Zone would inhibit the growth of deep roots, D~ext~ in the Discharge Zone is set at 0.5 m in basecase conditions.

The first set of Native Vegetation scenarios (category NatVeg) retains the relevant assumptions of the basecase regarding the native vegetation k~c~ and ET from the aquifer; namely, native vegetation is assigned a uniform k~c~ of 0.6 and ET from groundwater occurs only in the Discharge Zone. The second set of Native Vegetation scenarios (category NatVegET) alters both of these assumptions (see next section).

For all following scenarios, native vegetation is assigned a k~c~ of 0.6, ET from groundwater occurs only in the Discharge Zone, and D~ext~ in the Discharge Zone is 0.5 m.

* **natveg_outside_adj** -- Cease irrigation serving *all fields outside the Adjudicated Zone* and convert affected acreage to native vegetation. 

* **natveg_gwmixed_outside_adj** -- Cease irrigation serving fields *outside the Adjudicated Zone, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 

* **natveg_inside_adj** -- Cease irrigation serving *all fields inside the Adjudicated Zone* and convert affected acreage to native vegetation.  

* **natveg_gwmixed_inside_adj** -- Cease irrigation serving fields *inside the Adjudicated Zone, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 

* **natveg_all** -- Cease irrigation serving *all irrigated fields in the SVIHM model* and convert affected acreage to native vegetation.

* **natveg_gwmixed_all** -- Cease irrigation serving *all irrigated fields in the SVIHM model, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 


## 3.2 Higher-ET, deeper-rooted native vegetation replaces agricultural acreage, various configurations (NatVegET)

As mentioned in the previous section, the first set of Native Vegetation scenarios (category NatVeg) retains the relevant assumptions of the basecase regarding the native vegetation k~c~ and ET from the aquifer; namely, native vegetation is assigned a uniform k~c~ of 0.6 and ET from groundwater occurs only in the Discharge Zone. However, this set of assumptions constrained the native vegetation cover in ways that may be unrealistic. A vegetation type that may fit the assumptions described above is a chaparral ecosystem, with shallow roots and low water demand relative to cultivated crops.

In the second set of Native Vegetation scenarios (category NatVegET), the authors of this study explored the potential hydrologic effects of establishment of a higher-ET, deeper-rooted native vegetation cover, such as a 10+ year old forest. To simulate these conditions, it was assumed that native vegetation had a maximum rooting depth (D~ext~) of up to 4.5 m deep, and native vegetation was assigned a crop coefficient of 1.0 (similar to irrigated crops). However, in the Discharge Zone, shallow groundwater was still assumed to inhibit deeper root growth, and so all areas with in the Discharge Zone retained a D~ext~ of 0.5 m.

For all following scenarios, native vegetation is assigned a k~c~ of 1.0, ET from groundwater occurs in the Discharge Zone and under all native vegetation, D~ext~ in the Discharge Zone is 0.5 m, and D~ext~ outside the Discharge Zone (beneath native vegetation) is 4.5 m.


* **natveg_outside_adj_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving *all fields outside the Adjudicated Zone* and convert affected acreage to native vegetation. 

* **natveg_gwmixed_outside_adj_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving fields *outside the Adjudicated Zone, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 

* **natveg_inside_adj_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving *all fields inside the Adjudicated Zone* and convert affected acreage to native vegetation.  

* **natveg_gwmixed_inside_adj_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving fields *inside the Adjudicated Zone, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 

* **natveg_all_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving *all irrigated fields in the SVIHM model* and convert affected acreage to native vegetation.

* **natveg_gwmixed_all_et_check_1.0nvkc_4.5m_ext** -- Cease irrigation serving *all irrigated fields in the SVIHM model, which have a groundwater or mixed groundwater and surface water irrigation source* and convert affected acreage to native vegetation. 


<!-- # 4. Other scenarios not related to management option selection -->

<!-- ## 4.1 Climate inputs -->

<!-- Future climate change scenarios (ET, precip, and streamflow) -->


<!-- ## 4.2 Physical streambed alteration  -->

<!-- BDAs (debatable effect) -->

<!-- * **bdas_tribs**  -->
<!-- * **bdas_all_streams**  -->
<!-- * **bdas_scott_r_mainstem**  -->

<!-- *elaborate* -->

<!-- ## 4.3 Scenarios discussed but not simulated -->

<!-- The following scenarios were discussed during GSP development but were not simulated. -->

<!-- The below proposed scenarios that were not simulated: -->

<!-- * Upper watershed vegetation management -->
<!-- * Upper watershed logging road removal/vegetation restoration -->
<!-- * Pumping liquid water uphill in winter and spraying it at elevation during sub-freezing periods, where it would be stored in an ice field -->

# References
