---
title: "Supplement D: Results with Hydrologic Benefit Function based on Chinook juvenile abundance"
author: "Claire Kouba"
date: "2026-02-13"
output: 
  bookdown::word_document2:
    fig_caption: yes
    number_sections: true
    global_numbering: true
bibliography: library.bib

---

```{r modelNamer, echo = FALSE, results = 'asis'}

# "Supplement B: Results with Hydrologic Benefit Function based on coho juvenile abundance"
# model_name = "coho_lasso_juv_on_hydro"

# "Supplement C: Results with Hydrologic Benefit Function based on coho juveniles per spawner"
# model_name = "coho_lasso_jps_on_hydro"

# "Supplement D: Results with Hydrologic Benefit Function based on Chinook juvenile abundance"
model_name = "chinook_lasso_juv_on_hydro"

# "Supplement E: Results with Hydrologic Benefit Function based on Chinook juveniles per spawner"
# model_name = "chinook_lasso_jps_on_hydro"


```

```{r setup_supp, echo = FALSE, warning=F, comment=F}

dev_mode = T
if(dev_mode){
  knitr::opts_chunk$set(echo = TRUE)
  
  suppressPackageStartupMessages({
    library(sf)
    library(terra)
    library(geodata)
    library(rpostgis)
    library(tidyr)
    library(plyr) #ssurgo and climate data
    library(data.table) #fread function
    library(dataRetrieval) # USGS flow data
    library(lubridate)
    library(httr)
    library(stringr)
    library(rstudioapi)
    library(dataRetrieval) # for usgs data
    library(data.table) #month function
    library(zoo) # rolling means
    library(rgl) # plot3d
    library(MARSS)
    #spatial data libraries
    library(RANN)
    library(tmap)
    library(RColorBrewer)
    library(flextable)
    library(scales) # for alpha() color function
    library(ggthemes) # for colorblind_pal()
    library(colorspace) # for sequential_hcl()
    library(ggpubr) # to use get_palette()
    library(here)
  }
  )
  ms_dir = here::here()
  data_dir = file.path(here::here(), "Data")
  fig_dir = file.path(ms_dir, "Graphics and Supplements")
  load(file.path(ms_dir, "manuscript_data.RData")) # Load data layers/tables
  
  source(file.path(ms_dir,"02_Figure_Functions.R"))   # Load figure functions
  # print figures for slides, yes or no?
  save_figs_here = fig_dir#"C:/Users/Claire/Dropbox/Documents/UCD/Presentations_Talks_Workshops_miniprojects/2022.02.17 Peer Group presentation Ch. 1/figs"
  print_for_slides=T; fig_i = 11
  save_landuse_figure = F
  
}
```
# Methodological details of Hydrologic Benefit function formulations (Supplements B-E)

The Hydrologic Benefit (HB) function is intended to estimate the ecological value (in units of fish production) provided to fish by specific flow values (e.g., timing of fall flow increase; magnitude of dry season flow). This is expressed in conceptual mathematical terms as $fish = b + m_1 * flow_1 + m_1*flow_2~{\cdots}$; see methods in  @KoubaEtAlWatershedSpecific2025 for more details. It's important to note that, in this dataset, only 40-60% of variation in fish outcome was explained in models based on flow data, and the HB function does not constitute a prediction of the number of fish outmigrating in a given year. 

Specifically, the results below are based on statistical models, using LASSO regression, of reproductive output for two species of salmon (coho and Chinook) in the Scott River stream system. Reproductive output is expressed in units of outmigrating juvenile abundance and normalized juveniles-per-adult spawner; two reproductive units for two species yields the four HB function formulations. 

While MARSS models were used as a tool for identifying important functional flows in [@KoubaEtAlWatershedSpecific2025], they are unsuitable for this hydrologic scenario outcome forecasting purpose, because they are autoregressive in nature: the forecasting of future fish outcomes requires historical observations of fish outcomes. Therefore, to use MARSS models in estimating the ecological effects of flow changes under different management scenarios, it would be necessary to have historical fish records collected in an alternate reality under the different management scenarios, which is not possible. Similarly, LASSO models which included spawner abundance as a predictor were also not included in this exercise. 

Consequently, only LASSO models dependent on hydrologic variables were evaluated as potential tools to forecast the outcomes of flow change on fish reproduction. 

For each species, across all statistical methods, the coefficients for any single predictor were consistently negative or positive; e.g., more coho reproduction was consistently associated with *earlier* fall reconnection. 


```{r coefTabCaptionMaker, echo = F}

if(dev_mode){
  coefs_tab = read.csv(file.path(data_dir, "kouba_et_al_2025_ensemble_coefs_chinook and coho.csv"))
  model_names = colnames(coefs_tab)[2:13]
capmaker_all_models = model_name_to_coefs_df(model_names)
}

```

# HB coefficients, component contributions, and annual values


```{r coefTab01, echo = FALSE, results = 'asis'}

coef_tab = coefs_tab[,c("Predictor",model_name, "Greater.Hydrologic.Benefit.value.associated.with")]
coef_tab = coef_tab[coef_tab[,model_name] != 0,]

#Create caption and column labels
attr = model_name_to_coefs_df(model_name)
caption_text = results8_cap_maker(attr)
column_labels = c("Predictor", "Value", "Greater Hydrologic Benefit value associated with")

colnames(coef_tab) = column_labels

hb_ft = flextable(data = coef_tab)
hb_ft = set_caption(hb_ft, caption = caption_text)
hb_ft = flextable::width(hb_ft, j = 1,  width = 1.5)
hb_ft = flextable::width(hb_ft, j= 2,  width = 1)
hb_ft = flextable::width(hb_ft, j=3,  width = 3)
hb_ft

```


```{r farmerFishTradeoffTable01, echo = F, warning =F}

scenario_tab_init = get_scenario_tab_init()
scenario_tab_init = scenario_tab_init[scenario_tab_init$scenario_category!="Attr",]
coef_tab = coef_tab[, c("Predictor", "Value")]

# Calculate mean of 35 years of hydrologic benefit value and ET
obj_fn_results = get_HBF_and_crop_ET_columns(scenario_tab = scenario_tab_init,
                                        lasso_weights = coef_tab)

scenario_tab = obj_fn_results$scenario_tab
obj_fn_wy = obj_fn_results$obj_fn_results_by_wy

# replace transformed data with straight abundance data for better interpretability

# for scenario tab
scenario_tab$hb_untrans_mean = 10^scenario_tab$HBF_mean
scenario_tab$hb_untrans_stdev = 10^scenario_tab$HBF_stdev

colnames(scenario_tab) = gsub(pattern = "HBF",
                              replacement = "HBF_log10",
                              x = colnames(scenario_tab))
colnames(scenario_tab) = gsub(pattern = "hb_untrans",
                              replacement = "HBF",
                              x = colnames(scenario_tab))

# for obj_fn_wy
obj_fn_wy$hbf_total_untrans = 10^obj_fn_wy$hbf_total
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hbf_total"] = "hbf_total_log10"
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hbf_total_untrans"] = "hbf_total"

```


```{r scenario_tab_plot_columns01, echo = F, warning=F}

# Plot specs
scen_cat_tab = get_scen_cat_tab()
scen_cat_tab = add_plot_cols_to_scen_cat_tab(scen_cat_tab_1 = scen_cat_tab)

# Assign category featurs to full scenario table
scenario_tab$color = scen_cat_tab$color[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$feas_cat = scen_cat_tab$feas_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$mgmt_cat = scen_cat_tab$mgmt_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$symbol = scen_cat_tab$symbol[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab = add_pareto_col_to_scen_tab(scenario_tab = scenario_tab)

```

```{r benefitByWY01, echo = F, fig.cap="\\label{fig:benefitByWY} Comparison of annual performance in two objective functions (Hydrologic Benefit and ET, representing environmental and agricultural objectives, respectively) for brood years 1991-2025. The scenarios pictured are basecase (black circles) and a scenario from each of three categories: Enhanced Recharge, Crop Change, and Native Vegetation (Low ET). Dry-type water years (vertical bars) are 1991, ’92, ’94, 2001, ’09, ’13, ’14, ’18, and ’20-’22." , warning=F}

# Plot prep
plot_scenarios1 = c("basecase",  # "maxMAR_fields2024", "grain_12k",# "natveg_low_8k",
                    "alf_curtail_08.15","alf_curtail_07.31","alf_curtail_07.15")
scen_cols = get_palette(c("black","goldenrod","green4","dodgerblue3"),k=4)
y_val_x_axis_label = model_name_to_y_val_label(model_name)

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 7.5, units = "in", res = 300)

  par(mfcol = c(2,1))
  plot_obj_fn_by_wy(plot_scenarios = plot_scenarios1,
                    cols = scen_cols,
                    panel_labels = c("A","B"),
                    y_val_label = y_val_x_axis_label)
  dev.off()
}

knitr::include_graphics(fig_path)
fig_i = fig_i + 1

```

```{r HBComponents, echo = F, fig.cap="\\label{fig:hbComponents} For the basecase scenario, contributions to annual Hydrologic Benefit function values, by functional flow metric (see Table 1 above; intercept values not shown)." , warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  plot_obj = hbf_comp_fig(obj_fn_wy, 
               scen_id = "basecase",#"alf_curtail_07.31",#"basecase",
               y_val_label = y_val_x_axis_label)
  fig_path
  ggsave(filename = fig_path, plot = plot_obj, width = 8, height = 6, units = "in"
    )
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1

```

\newpage
# Trade-offs and efficiency

```{r farmerFishTradeoffFigure01, echo = F, fig.cap="\\label{fig:farmerFishTradeoffFigure} Environmental and agricultural objective function performance (in terms of Hydrologic Benefit, or HB, value; and crop ET volume) for basecase (black circle in upper left) and 25 management scenarios. Symbol x and y coordinates are the average of 35 annual HB and ET values for each scenario over the full model period (water years 1991-2025); for exampe time series of annual values, see previous figure." , fig.height=8, fig.width=7, warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_fig(scenario_tab = scenario_tab,
                         y_val_label = y_val_x_axis_label,
                         plot_title = "Environmental vs Agricultural Benefit \n Mean of annual values, 1991-2025")
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1

#Make blank axes for slides
if(print_for_slides==T){png(filename = file.path(save_figs_here, paste0("Figure ",fig_i,".png")), width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_blank(scenario_tab = scenario_tab,
                           y_val_label = y_val_x_axis_label)
  fig_i = fig_i +1; dev.off()}

```


```{r efficiency01, echo = F, fig.cap="\\label{fig:efficiency} Scenario trade-off efficiencies, or average Hydrologic Benefit value (a proxy for environmental benefit) gained per crop ET value lost (a proxy for farmer cost) (normalized to average basecase HB and ET values, respectively), in all 35 years. Some scenarios are excluded if they have no or negative ET loss, producing infinite or negative efficiency values." , warning=F}


bc_sel = scenario_tab$scenario_id=="basecase"
hb_bc_mean = scenario_tab$HBF_mean[bc_sel]
et_bc_mean = scenario_tab$et_mean[bc_sel]
scenario_tab$hb_gained = (scenario_tab$HBF_mean - hb_bc_mean)/hb_bc_mean
scenario_tab$et_lost = (scenario_tab$et_mean - et_bc_mean)/et_bc_mean *-1

scenario_tab$hb_gained_to_et_lost = round(scenario_tab$hb_gained / scenario_tab$et_lost, 2)

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 6, units = "in", res = 300)
  tradeoff_efficiency_fig_nov2025(scenario_tab = scenario_tab)
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1




```




# Three objectives and Pareto-optimal set


```{r farmerFishInfraFigure3DSupp, echo = F, fig.align='center', fig.cap="\\label{fig:farmerFishInfraFigure3D} A three-dimensional scatter plot of the objective function values for basecase and 24 management scenarios, averaged over 35 water years (1991-2025). Larger spheres are shown for the Pareto-optimal set of scenarios. Scenario colors correspond to legend in Supplement Figures 3 and 4. The corner corresponding to optimal values of all objective functions (the 'best' corner) is highlighted with a red arrow." , out.width="80%", warning=F}


save_3d_plot = F

if(save_3d_plot == T){
  # need to run pareto_set_calc R chunk first
plot3d(x = scenario_tab$HBF_mean/1000000, 
       y = scenario_tab$et_mean / (-1 * 10^6), 
       z = scenario_tab$feas_cat,
       type = "s", col = scenario_tab$color, 
       radius = scenario_tab$size_3d,
       xlab = "HB/10^6",#"Hyd. Ben Value",
       ylab = "ET vol.",
       zlab = "Infrastructure")
  
}
#This only works to manually orient it and take a screen shot; do this if you need to redo this graph

if(model_name == "coho_lasso_juv_on_hydro"){
  scatter3d_filename = file.path(fig_dir, "hbf et infra plot3d_2026.03.06_coho juv.jpg")
}
if(model_name == "coho_lasso_jps_on_hydro"){
  scatter3d_filename = file.path(fig_dir, "hbf et infra plot3d_2026.03.06_coho jps.jpg")
}
if(model_name == "chinook_lasso_juv_on_hydro"){
  scatter3d_filename = file.path(fig_dir, "hbf et infra plot3d_2026.03.06_chinook juv.jpg")
}
if(model_name == "chinook_lasso_jps_on_hydro"){
  scatter3d_filename = file.path(fig_dir, "hbf et infra plot3d_2026.03.06_chinook jps.jpg")
}

knitr::include_graphics(scatter3d_filename)

```




\newpage

# Scenario Results Table

```{r scenarioResultsTab01, echo = FALSE, results = 'asis'}

# subset desired columns
scen_tab = scenario_tab[,c("scenario_id", "scenario_category", "feas_cat", 
                           "HBF_mean", "et_mean", "pareto_opt",
                           "hb_gained_to_et_lost")]
if(max(scen_tab$HBF_mean) > 2000){
  scen_tab$HBF_mean = round(scen_tab$HBF_mean / 1000, 2)
  mean_HB_text = "Mean HB / 1000"
} else {
  scen_tab$HBF_mean = round(scen_tab$HBF_mean, 2)
  mean_HB_text = "Mean HB"
}

scen_tab$et_mean = round(scen_tab$et_mean)
# remove the negative efficiencies - that means ET was gained, not lost, so it doesn't apply
scen_tab$hb_gained_to_et_lost = round(scen_tab$hb_gained_to_et_lost, 2)
scen_tab$hb_gained_to_et_lost[scenario_tab$hb_gained <=0 | 
                                scenario_tab$et_lost <=0 |
                                !is.finite(scen_tab$hb_gained_to_et_lost) |
                                scen_tab$scenario_id == "basecase" ] = "--"
scen_tab$et_mean = round(scenario_tab$et_mean * (-1 / 10^6), 2)

# Exclude attribution scenarios
scen_tab = scen_tab[scen_tab$scenario_category != "Attr",]

#Create caption and column labels
caption_text = "Scenario attributes for 25 scenarios. Values in the Mean Hydrologic Benefit (HB) and Mean ET columns are averages of 35 annual values for each scenario. Efficiency not calculated for scenarios with no average ET loss."
column_labels = c("Scenario ID", "Category", "Feas.", mean_HB_text, "Mean ET",  "Pareto", "Efficiency")

scen_tab_for_ft = scen_tab
colnames(scen_tab_for_ft) = column_labels

st_ft = flextable(scen_tab_for_ft)
st_ft = set_caption(st_ft, caption = caption_text)

st_ft = flextable::fontsize(st_ft,j=1,  size = 10)
st_ft = flextable::fontsize(st_ft,j=2,  size = 10)
st_ft = flextable::fontsize(st_ft,j=3,  size = 10)
st_ft = flextable::fontsize(st_ft,j=4,  size = 10)
st_ft = flextable::fontsize(st_ft,j=5,  size = 10)
st_ft = flextable::fontsize(st_ft,j=6,  size = 10)
st_ft = flextable::fontsize(st_ft,j=7,  size = 10)

# st_ft = flextable::rotate(st_ft, rotation="btlr",j=1:8)
st_ft = flextable::width(st_ft,j=1,  width = 1.7)
st_ft = flextable::width(st_ft,j=2,  width = 1)
st_ft = flextable::width(st_ft,j=3,  width = .7)
st_ft = flextable::width(st_ft,j=4,  width = .7)
st_ft = flextable::width(st_ft,j=5,  width = .7)
st_ft = flextable::width(st_ft,j=6,  width = .8)
st_ft = flextable::width(st_ft,j=7,  width = .8)

st_ft


``` 


\newpage
# Functional Flow Metrics 

Annual values of the functional flow metrics highlighted in this Hydrologic Benefit function are tabulated below. The basecase and alfalfa irrigation scenarios are used as examples; for a graph of each scenario's impact on simulated Fort Jones daily average flow, see Supplement E, Figure 2. 

\newpage

```{r funcFlowValsTab01, echo = FALSE, results = 'asis'}

#Initialize things for all tables
ff_tab_list = ff_tab_builder(scen_ids = plot_scenarios1, coef_tab = coef_tab)
predictors = coef_tab$Predictor[coef_tab$Predictor!="Intercept"]

all_func_flows = colnames(ffs_scenarios[['basecase']])[colnames(ffs_scenarios[['basecase']]) != "Brood_Year"]
ff_digits = data.frame(func_flow = all_func_flows, digits = 0)
ff_digits$digits[ff_digits$func_flow %in% c("s1_SP_ROC","s1_SP_ROC_Max")] = 3
ff_digits = rbind(ff_digits,
                  data.frame(func_flow = "f1_recon_120", digits = 0))
col_width_uni = 6.5/(length(plot_scenarios1)+1)

# Make table for 1st predictor
ff_i = predictors[1]

#Create caption and column labels
  caption_text = paste0("For functional flow ",ff_i,": Annual functional flow values, for example scenarios, based on simulated Fort Jones daily average flow, which influence Hydrologic Benefit Function results.")
fftab = ff_tab_list[[1]]
fft_ft = flextable(data = fftab)
fft_ft = set_caption(fft_ft, caption = caption_text)
fft_ft = colformat_double(fft_ft, j=2:ncol(fftab), 
                          digits = ff_digits$digits[ff_digits$func_flow==ff_i])
fft_ft = colformat_num(fft_ft, j=1, big.mark="")
fft_ft = width(fft_ft, width=col_width_uni)
fft_ft

```

\newpage

```{r funcFlowValsTab02, echo = FALSE, results = 'asis'}

if(length(predictors)>1){
  ff_i = predictors[2]
  #Create caption and column labels
  caption_text = paste0("For functional flow ",ff_i,": Annual functional flow values, for example scenarios, based on simulated Fort Jones daily average flow, which influence Hydrologic Benefit Function results.")
  fftab = ff_tab_list[[2]]
  fft_ft = flextable(data = fftab)
  fft_ft = set_caption(fft_ft, caption = caption_text)
  fft_ft = colformat_double(fft_ft, j=2:ncol(fftab), 
                          digits = ff_digits$digits[ff_digits$func_flow==ff_i])
  fft_ft = colformat_num(fft_ft, j=1, big.mark="")
  fft_ft = width(fft_ft, width=col_width_uni)
  fft_ft
}

```

\newpage

```{r funcFlowValsTab03, echo = FALSE, results = 'asis'}

if(length(predictors)>2){
  ff_i = predictors[3]
  #Create caption and column labels
    caption_text = paste0("For functional flow ",ff_i,": Annual functional flow values, for example scenarios, based on simulated Fort Jones daily average flow, which influence Hydrologic Benefit Function results.")
  fftab = ff_tab_list[[3]]
  fft_ft = flextable(data = fftab)
  fft_ft = set_caption(fft_ft, caption = caption_text)
  fft_ft = colformat_double(fft_ft, j=2:ncol(fftab), 
                            digits = ff_digits$digits[ff_digits$func_flow==ff_i])
  fft_ft = colformat_num(fft_ft, j=1, big.mark="")
  fft_ft = width(fft_ft, width=col_width_uni)
  fft_ft
}

```


\newpage

```{r funcFlowValsTab04, echo = FALSE, results = 'asis'}

if(length(predictors)>3){
  ff_i = predictors[4]
  #Create caption and column labels
  caption_text = paste0("For functional flow ",ff_i,": Annual functional flow values, based on simulated Fort Jones daily average flow, influencing Hydrologic Benefit Function results.")
  fftab = ff_tab_list[[4]]
  fft_ft = flextable(data = fftab)
  fft_ft = set_caption(fft_ft, caption = caption_text)
  fft_ft = width(fft_ft, width=col_width_uni)
  fft_ft
}

```

