---
title: "Multi-objective assessment of a stakeholder-defined portfolio of groundwater and stream management actions in an agricultural basin"
author: "Claire Kouba"
date: "2025-11-12"
output: 
  bookdown::word_document2:
    fig_caption: yes
    number_sections: true
    global_numbering: true
bibliography: library.bib

---

# Fish-Farm Results of 4 HB function formulations


While MARSS models were used as a tool for identifying important functional flows in [@KoubaEtAlWatershedSpecific2025], they are unsuitable for this hydrologic scenario outcome forecasting purpose, because they are autoregressive in nature: the forecasting of future fish outcomes requires historical observations of fish outcomes. Therefore, to use MARSS models in estimating the ecological effects of flow changes under different management scenarios, it would be necessary to have historical fish records collected under the different management scenarios, which are not possible. 

Consequently, only LASSO models dependent on hydrologic variables (and not including spawner abundance as a predictor) were evaluated as potential tools to forecast the outcomes of flow change on fish reproduction.


```{r setup_supp, echo = FALSE, warning=F, comment=F}

dev_mode = T#F
if(dev_mode){
  knitr::opts_chunk$set(echo = TRUE)
  
  suppressPackageStartupMessages({
    library(sf)
    library(terra)
    library(geodata)
    library(rpostgis)
    library(tidyr)
    library(plyr) #ssurgo and climate data
    library(data.table) #fread function
    library(dataRetrieval) # USGS flow data
    library(lubridate)
    library(httr)
    library(stringr)
    library(rstudioapi)
    library(dataRetrieval) # for usgs data
    library(data.table) #month function
    library(zoo) # rolling means
    library(rgl) # plot3d
    library(MARSS)
    #spatial data libraries
    library(RANN)
    library(tmap)
    library(RColorBrewer)
    library(flextable)
    library(scales) # for alpha() color function
    library(ggthemes) # for colorblind_pal()
    library(colorspace) # for sequential_hcl()
    library(ggpubr) # to use get_palette()
    library(here)
  }
  )
  ms_dir = here::here()
  data_dir = file.path(here::here(), "Data")
  fig_dir = file.path(ms_dir, "Graphics and Supplements")
  load(file.path(ms_dir, "manuscript_data.RData")) # Load data layers/tables
  
  source(file.path(ms_dir,"02_Figure_Functions.R"))   # Load figure functions
  # print figures for slides, yes or no?
  save_figs_here = fig_dir#"C:/Users/Claire/Dropbox/Documents/UCD/Presentations_Talks_Workshops_miniprojects/2022.02.17 Peer Group presentation Ch. 1/figs"
  print_for_slides=T; fig_i = 1
  save_landuse_figure = F
  
}
```

```{r coefTabCaptionMaker, echo = F}

if(dev_mode){
  coefs_tab = read.csv(file.path(data_dir, "kouba_et_al_2025_ensemble_coefs_chinook and coho.csv"))
}
model_names = colnames(coefs_tab)[2:13]
capmaker_all_models = model_name_to_coefs_df(model_names)

```



## Models of coho reproductive success

### LASSO model, juvenile abundance

```{r coefTab01, echo = FALSE, results = 'asis'}

# model_name = "chinook_lasso_jps_on_hydro"
# model_name = "chinook_lasso_juv_on_hydro"
# model_name = "coho_lasso_jps_on_hydro"
model_name = "coho_lasso_juv_on_hydro"


coef_tab = coefs_tab[,c("Predictor",model_name, "Greater.Hydrologic.Benefit.value.associated.with")]
coef_tab = coef_tab[coef_tab[,model_name] != 0,]

#Create caption and column labels
attr = model_name_to_coefs_df(model_name)
caption_text = results8_cap_maker(attr)
column_labels = c("Predictor", "Value", "Greater Hydrologic Benefit value associated with")

colnames(coef_tab) = column_labels

hb_ft = flextable(data = coef_tab)
hb_ft = set_caption(hb_ft, caption = caption_text)
hb_ft = flextable::width(hb_ft, j = 1,  width = 1.5)
hb_ft = flextable::width(hb_ft, j= 2,  width = 1)
hb_ft = flextable::width(hb_ft, j=3,  width = 3)
hb_ft

```


```{r farmerFishTradeoffTable01, echo = F, warning =F}

scenario_tab_init = get_scenario_tab_init()
scenario_tab_init = scenario_tab_init[scenario_tab_init$scenario_category!="Attr",]
coef_tab = coef_tab[, c("Predictor", "Value")]

# Calculate mean of 35 years of hydrologic benefit value and ET
obj_fn_results = get_HBF_and_crop_ET_columns(scenario_tab = scenario_tab_init,
                                        lasso_weights = coef_tab)

scenario_tab = obj_fn_results$scenario_tab
obj_fn_wy = obj_fn_results$obj_fn_results_by_wy

# replace transformed data with straight abundance data for better interpretability

# for scenario tab
scenario_tab$hb_untrans_mean = 10^scenario_tab$HBF_mean
scenario_tab$hb_untrans_stdev = 10^scenario_tab$HBF_stdev

colnames(scenario_tab) = gsub(pattern = "HBF",
                              replacement = "HBF_log10",
                              x = colnames(scenario_tab))
colnames(scenario_tab) = gsub(pattern = "hb_untrans",
                              replacement = "HBF",
                              x = colnames(scenario_tab))

# for obj_fn_wy
obj_fn_wy$hb_val_untrans = 10^obj_fn_wy$hb_val
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hb_val"] = "hb_val_log10"
colnames(obj_fn_wy)[colnames(obj_fn_wy)=="hb_val_untrans"] = "hb_val"

# # Add ET columns from ET table
# scenario_tab$et_mean = scen_tab_et$et_mean[match(scenario_tab$scenario_id,
#                                                  scen_tab_et$scenario_id)]
# scenario_tab$et_stdev = scen_tab_et$et_stdev[match(scenario_tab$scenario_id,
#                                                  scen_tab_et$scenario_id)]
```


```{r scenario_tab_plot_columns01, echo = F, warning=F}

# Plot specs
scen_cat_tab = get_scen_cat_tab()
scen_cat_tab = add_plot_cols_to_scen_cat_tab(scen_cat_tab_1 = scen_cat_tab)

# Assign category featurs to full scenario table
scenario_tab$color = scen_cat_tab$color[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$feas_cat = scen_cat_tab$feas_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$mgmt_cat = scen_cat_tab$mgmt_cat[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab$symbol = scen_cat_tab$symbol[match(scenario_tab$scenario_category,
                                                    scen_cat_tab$category)]
scenario_tab = add_pareto_col_to_scen_tab(scenario_tab = scenario_tab)

```



```{r benefitByWY01, echo = F, fig.cap="\\label{fig:benefitByWY} Comparison of annual performance in two objective functions (HB and ET, representing environmental and agricultural objectives, respectively), in the basecase (black circles) and management scenarios (diamond, cross or x symbols) for water years 1991-2025. Panels A and B show comparisons between basecase and one example scenario in each of three categories: Enhanced Recharge, Low Flow Diversion Limits, and Curtailment. Panels A and B show comparisons between basecase and different cutoff dates within the Alfalfa Irrigation category: July 15th, July 31st and August 15th. Dry-type water years (vertical bars) are 1991, '92, '94, 2001, '09, '13, '14, '18, and '20-'22." , fig.height=8.3, fig.width=7, warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 8, units = "in", res = 300)

  # plot_scenarios = scenario_tab$scenario_id
  plot_scenarios1 = c("basecase", "alf_curtail_08.15", "alf_curtail_07.31",
                      "alf_curtail_07.15")
  plot_scenarios2 = c("basecase", "maxMAR_fields2024", "eflows25_div_lims",
                      "grain_12k")

  par(mfcol = c(4,1))
  plot_obj_fn_by_wy(plot_scenarios = plot_scenarios1,
                    cols = get_palette(c("black","dodgerblue",
                                         "green4","darkgoldenrod"),
                                       length(plot_scenarios1)),
                    panel_labels = c("A","B"))
  plot_obj_fn_by_wy(plot_scenarios = plot_scenarios2,
                    cols = sequential_hcl(n=length(plot_scenarios1),
                                          palette = "Red-Blue"),
                    panel_labels = c("C","D"))
  dev.off()
}

knitr::include_graphics(fig_path)
fig_i = fig_i + 1

```

```{r farmerFishTradeoffFigure01, echo = F, fig.cap="\\label{fig:farmerFishTradeoffFigure} Environmental and agricultural objective function performance (in terms of Hydrologic Benefit, or HB, value; and crop ET volume) for basecase (black circle in upper left) and 39 management scenarios. Symbol x and y coordinates are the average of 28 annual HB and ET values for each scenario over the full model period (water years 1991-2018). Whiskers show standard error of the 28 years of annual HB values; standard errors in the direction of the y-axis are smaller than the height covered by each symbol." , fig.height=8, fig.width=7, warning=F}

fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_fig(scenario_tab = scenario_tab,
                         plot_title = "Environmental vs Agricultural Benefit \n Mean of annual values, 1991-2025")
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1

#Make blank axes for slides
if(print_for_slides==T){png(filename = file.path(save_figs_here, paste0("Figure ",fig_i,".png")), width = 8, height = 8, units = "in", res = 300)
  farm_fish_tradeoff_blank(scenario_tab = scenario_tab)
  fig_i = fig_i +1; dev.off()}

```


```{r efficiency, echo = F, fig.cap="\\label{fig:efficiency} Scenario trade-off efficiencies, or average HB value (a proxy for environmental benefit) gained per crop ET value lost (a proxy for farmer cost) (normalized to average basecase HB and ET values, respectively), in all 35 years. Basecase, Reservoir scenarios and Enhanced Recharge scenarios are not included as they have on average no or negative ET loss, producing infinite or negative efficiency values." , warning=F}



fig_path = file.path(save_figs_here, paste0("Figure ",fig_i,".png"))
if(print_for_slides==T){
  png(filename = fig_path, width = 8, height = 6, units = "in", res = 300)
  tradeoff_efficiency_fig_nov2025(scenario_tab = scenario_tab)
  dev.off()
}
knitr::include_graphics(fig_path)
fig_i = fig_i + 1




```

<!-- ### MARSS model, juvenile abundance -->



<!-- ```{r MARSS_models_single_covar, echo = F} -->
<!-- # Specify MARSS model ------------------------------------------------------------------ -->

<!-- model_list_file = "single cov MARSS models.rds" -->
<!-- coef_tab_file = "single cov marss coefs.csv" -->
<!-- aicc_tab_file = "single cov MARSS AICc tab.csv" -->

<!-- if(file.exists(file.path(data_dir, model_list_file))){ -->
<!--   list_of_mods = readRDS(file.path(data_dir, model_list_file)) -->
<!--   aicc_tab = read.csv(file.path(data_dir, aicc_tab_file)) -->
<!--   coef_tab = read.csv(file.path(data_dir, coef_tab_file)) -->

<!-- } else { -->
<!--   print("Error: MARSS files missing. Retrieve MARSS files from Kouba et al 2025 repository.") -->
<!-- } -->


<!-- ## The code in the functions called below follows the R script -->
<!-- ## Baruch_2023_Putah_Creek_MARSS.R, published in support of: -->
<!-- ## Baruch et al., 2024. "Mimicking Functional Elements of the Natural Flow Regime -->
<!-- ## Promotes Native Fish Recovery in a Regulated River." https://doi.org/10.1002/eap.3013. -->


<!-- ``` -->



<!-- ```{r MARSSSingleCovarTableCoho, echo = F, warning =F, message = F, results = 'asis'} -->

<!-- yvlt = kouba_2025_y_val_label_tab() -->

<!-- y_val = "coho_smolt_abun_est" -->
<!-- # y_val = "coho_smolt_per_fem" -->
<!-- # y_val = "chinook_juvenile_abundance" -->
<!-- # y_val = "chinook_juv_per_adult" -->

<!-- aicc_order = order(aicc_tab[,y_val]) -->
<!-- # view predictors in order of AICc value (smallest first) -->
<!-- results_y = data.frame(pred = aicc_tab[aicc_order, "pred"]) -->
<!-- results_y$coef = round(coef_tab[aicc_order, y_val], 3) -->
<!-- results_y$aicc = round(aicc_tab[aicc_order, y_val], 2) -->
<!-- colnames(results_y) = c("Covariate", "Coefficient", "AICc") -->
<!-- # make table for display -->
<!-- results_y_ft = results_y -->
<!-- didnt_model = is.na(results_y_ft$AICc) -->
<!-- results_y_ft$AICc[didnt_model] = "--"; results_y_ft$Coefficient[didnt_model] = "--" -->

<!-- caption_text = paste0("Each row corresponds to a MARSS model predicting the time series of ", yvlt$y_val_title[yvlt$y_val==y_val], " observations using itself (up to time t) and one hydrologic metric covariate. Coefficient sign and value indicate the direction and strength of the influence of the hydrologic metric; i.e., a negative coefficient for hydrologic metric f1_recon_120 indicates that an earlier first fall river reconnection (120 cfs) is associated with a greater ", yvlt$y_val_title[yvlt$y_val==y_val], " value. Models are listed in order from best (lowest AICc value) to worst.  Values marked with -- indicate that gaps in the time series for the hydrologic metric prevented the calculation of a model using that covariate.") -->

<!-- co_tab_ft = flextable(data=results_y_ft) -->
<!-- co_tab_ft = set_caption(co_tab_ft, caption = caption_text) -->
<!-- co_tab_ft = flextable::width(co_tab_ft,j=1,  width = 2) -->
<!-- co_tab_ft = flextable::width(co_tab_ft,j=2,  width = 1) -->
<!-- co_tab_ft = flextable::width(co_tab_ft,j=3,  width = 1) -->
<!-- co_tab_ft -->


<!-- ``` -->




<!-- ```{r MARSSSingleCovars, echo = F, warning = F, fig.height = 8, fig.cap = "Results of the three best single-hydrologic-covariate MARSS models to predict log-transformed Chinook and coho juvenile-per-spawners ratios with Z-scored hydrologic metrics.\\label{fig:MARSSSingleCovars}"} -->

<!-- # pull or read in metrics tabs from all scenarios -->

<!-- scenario_tab_init = get_scenario_tab_init() -->
<!-- scenario_tab_init = scenario_tab_init[scenario_tab_init$scenario_category!="Attr",] -->


<!-- # scenarios = scenario_tab_init$scenario_id -->
<!-- # i=1 -->


<!-- # # For MARSS auto-prediction: add y values to scenario-derived func flow metrics tab -->
<!-- metrics_tab_orig = read.csv(file.path(data_dir, "kouba_et_al_2025 Supplemental Table B. Flow Metrics by Brood Year.csv")) -->
<!-- # # add log-transformed eco y-value metrics -->
<!-- # ff_bc = ffs_scenarios[["basecase"]] -->
<!-- eco_y_vals = metrics_tab_orig[,c("brood_year",yvlt$y_val)] -->
<!-- eco_y_vals_log10 = eco_y_vals -->
<!-- eco_obs_yrs = list() -->
<!-- for(i in 1:nrow(yvlt)){ -->
<!--   y_val_i = yvlt$y_val[i] -->
<!--   # ff_bc[,y_val] = log10(eco_y_vals[match(ff_scen$Brood_Year, -->
<!--                                      # eco_y_vals$brood_year), y_val]) -->
<!--   eco_y_vals_log10[,y_val] = log10(eco_y_vals[ , y_val_i]) -->
<!--   metrics_tab_orig[,y_val_i] = log10(metrics_tab_orig[, y_val_i]) -->
<!--   non_nas = eco_y_vals_log10$brood_year[which(!is.na(eco_y_vals_log10[,y_val]))] -->
<!--   eco_obs_yrs[[y_val_i]] = min(non_nas):max(non_nas) -->
<!--   # names(eco_obs_yrs[[i]]) = y_val_i -->
<!-- } -->

<!-- pred_cols = colnames(metrics_tab_orig)[ -->
<!--   !(colnames(metrics_tab_orig) %in% c("brood_year", "smolt_year", yvlt$y_val))] -->
<!-- for(i in 1:length(pred_cols)){ -->
<!--   pred_col = pred_cols[i] -->
<!--   z_mean = mean(metrics_tab_orig[,pred_col], na.rm=T) -->
<!--   z_sd = sd(metrics_tab_orig[,pred_col], na.rm=T) -->
<!--   metrics_tab_orig[,pred_col] = (metrics_tab_orig[,pred_col] - z_mean)/z_sd -->
<!-- } -->

<!-- mod_num = 1# for plot -->
<!-- cov_i = results_y$Covariate[order(results_y$AICc)][mod_num] -->
<!-- mod_i = list_of_mods[[paste0(y_val, "__", cov_i)]] -->
<!-- yval_yrs = eco_obs_yrs[[y_val]] -->

<!-- # for troubleshooting marss y predictions -->
<!-- marss_predict_plot_single_covars(metrics_tab=metrics_tab_orig, -->
<!--                                  pred_yrs = yval_yrs, -->
<!--                                  y_val, -->
<!--                                  cov_i = cov_i, -->
<!--                                  mod_i = mod_i) -->

<!-- # Calculate mean of 35 years of hydrologic benefit value and ET -->

<!-- obj_fn_results = get_HBF_and_crop_ET_columns(scenario_tab = scenario_tab_init, -->
<!--                                              # metrics_tab = ff_scen, -->
<!--                                              y_val = y_val,  -->
<!--                                              cov_i = cov_i, -->
<!--                                              mod_i = mod_i, -->
<!--                                              yval_yrs = yval_yrs) -->

<!--   # calculate_MARSS_HBF(scenario_tab = scenario_tab_init, -->
<!--   #                                         metrics_tab = ff_scen, -->
<!--   #                                pred_yrs_i = pred_yrs_i, -->
<!--   #                                y_val = y_val,  -->
<!--   #                                cov_i = cov_i, -->
<!--   #                                mod_i = mod_i) -->

<!-- scenario_tab = obj_fn_results$scenario_tab -->
<!-- obj_fn_wy = obj_fn_results$obj_fn_results_by_wy -->






<!-- fig_path = file.path(save_figs_here, paste0("Figure S",fig_i,".png")) -->
<!-- if(save_imgs == TRUE){ -->
<!--   png(filename = fig_path, width = 7, height = 8, units = "in", res = res_dpi) -->

<!--   # par(mfrow = c(2,1)) -->


<!--   # marss_predict_plot_single_covars(metrics_tab = metrics_tab, -->
<!--   #                                  pred_yrs_i = pred_yrs_i_ch, -->
<!--   #                                  y_val = y_val_chinook, cov_i = cov_i_ch, -->
<!--   #                                  mod_i = mod_is_ch)   -->

<!--   dev.off() -->
<!-- } -->
<!-- if(include_figs==T){knitr::include_graphics(fig_path)} -->
<!-- fig_i = fig_i +1 -->


<!-- ``` -->
